---
title: "Data"
output: html_document
date: "2023-04-20"
---

```{r echo=FALSE, message=FALSE}
source(here::here("scripts/setup.R"))
```

```{r Loading and early cleaning, cache=TRUE}
#loading the data
recipes_raw <- read.csv(here("data/epi_r.csv"))

recipes <- recipes_raw%>% 
  clean_names() %>% 
  mutate(ID = 1:nrow(.)) %>%
  select(ID, everything())
```

# Categories from all the variables

TO DO:
- Merge "father_s_day" and "fathers_day" same for mother's day and for new year's and st patrick and valentines day, vermout and vermouth - DONE
- Merge cranberry and cranberry_sauce - DONE
- Check what leafy_green is and how many obs there is of it
- Same for "legume"
- Decide if we group beans together or leave them in vegetables - DONE
- What do we do with "meat", meatball and meatloaf. What about rabbit and for sausage or steak and venison - DONE
- Put nutmeg in spices or nuts? - DONE
- do we create a seeds category --> for example for "poppy" (put in spices for now) and for "seed" and sesame - DONE
- We should probably create a sauce category, and a "full_meal"
- where do we put tapioca and yuca - DONE
- do we put buttermilk in drinks or dairy?
- check if dorie_greenspa column exist or if it was just a typo without the N
- check how many observations with phyllo_puff_pastry_dough
- do we separate fish and seafood?


to CHECK:
- meat DONE
- legume - DONE
- drink and drinks
- fish or seafood - DONE
- herb - DONE
- nut - DONE
- fruit - DONE
- cheese - DONE
- dairy - DONE
- spice - DONE

NOTES:
final all_ingredients is 304 obs long, if you include the general categories in it, otherwise it's 293 (292 in vectors plus "egg")

ALWAYS CHECK THAT ingredients_df is 321 after the binary columns and general categories are added. 322 now with rating as well

```{r Creating Variable Categories}
us_states <- c("alabama", "alaska", "arizona", "california", "colorado", "connecticut", "florida", "georgia", "hawaii", "idaho", "illinois", "indiana", "iowa", "kansas", "kentucky", "louisiana", "maine", "maryland", "massachusetts", "michigan", "minnesota", "mississippi", "missouri", "nebraska", "new_hampshire", "new_jersey", "new_mexico", "new_york", "north_carolina", "ohio", "oklahoma", "oregon", "pennsylvania", "rhode_island", "south_carolina", "tennessee", "texas", "utah", "vermont", "virginia", "washington", "west_virginia", "wisconsin")

us_cities <- c("aspen", "atlanta", "beverly_hills", "boston","brooklyn", "buffalo", "cambridge", "chicago", "columbus", "costa_mesa", "dallas", "denver", "healdsburg", "hollywood", "houston", "kansas_city", "lancaster", "las_vegas", "london", "los_angeles", "louisville", "miami", "minneapolis", "new_orleans", "pacific_palisades", "paris", "pasadena", "pittsburgh", "portland", "providence", "san_francisco", "santa_monica", "seattle", "st_louis", "washington_d_c", "westwood", "yonkers")

countries <- c("australia", "bulgaria", "canada", "chile", "cuba", "dominican_republic", "egypt", "england", "france", "germany", "guam", "haiti", "ireland", "israel", "italy", "jamaica", "japan", "mexico", "mezcal", "peru", "philippines", "spain", "switzerland")

alcohol <- c("alcoholic", "amaretto", "beer", "bitters", "bourbon", "brandy", "calvados", "campari", "chambord", "champagne", "chartreuse", "cocktail", "cognac_armagnac", "creme_de_cacao", "digestif", "eau_de_vie", "fortified_wine", "frangelico", "gin", "grand_marnier", "grappa", "kahlua", "kirsch", "liqueur", "long_beach", "margarita", "marsala", "martini", "midori", "pernod", "port", "punch", "red_wine", "rum", "sake", "sangria", "scotch", "sherry", "sparkling_wine", "spirit", "spritzer", "tequila", "triple_sec", "vermouth", "vodka", "whiskey", "white_wine", "wine")

wtf <- c("bon_appetit", "bon_app_tit", "condiment_spread", "cr_me_de_cacao", "epi_ushg", "flaming_hot_summer", "frankenrecipe", "harpercollins", "house_garden", "no_meat_no_problem", "parade", "sandwich_theory", "self", "shower", "tested_improved", "windsor", "weelicious", "snack_week", "tailgating", "quick_and_healthy", "picnic", "kitchen_olympics", "house_cocktail", "hors_d_oeuvre", "frozen_dessert", "freezer_food", "edible_gift", "cookbook_critic", "cook_like_a_diner", "condiment", "cocktail_party", "camping", "buffet", "x30_days_of_groceries", "x_cakeweek", "x_wasteless", "x22_minute_meals", "x3_ingredient_recipes")

chef <- c("anthony_bourdain", "dorie_greenspan", "emeril_lagasse", "nancy_silverton", "suzanne_goin")

interesting <- c("advance_prep_required", "entertaining", "epi_loves_the_microwave", "friendsgiving", "game", "gourmet", "healthy", "high_fiber", "hot_drink", "kid_friendly", "kidney_friendly", "low_cal", "low_cholesterol", "microwave", "no_cook", "one_pot_meal", "oscars", "paleo", "peanut_free", "pescatarian", "poker_game_night", "potluck", "quick_easy", "sugar_conscious", "wheat_gluten_free", "cookbooks", "leftovers")

season <- c("winter", "spring", "summer", "fall")

celebrations <- c("anniversary", "back_to_school", "bastille_day", "birthday", "christmas", "christmas_eve", "cinco_de_mayo", "date", "diwali", "easter", "engagement_party", "family_reunion", "father_s_day", "fourth_of_july", "graduation", "halloween", "hanukkah", "kentucky_derby", "kwanzaa", "labor_day", "lunar_new_year", "mardi_gras", "mother_s_day", "new_year_s_day", "new_year_s_eve", "oktoberfest", "party", "passover", "persian_new_year", "purim", "ramadan", "rosh_hashanah_yom_kippur", "shavuot", "st_patrick_s_day", "sukkot", "super_bowl", "thanksgiving", "valentine_s_day", "wedding")

tools <- c("coffee_grinder", "double_boiler", "food_processor", "ice_cream_machine", "juicer", "mandoline", "mixer", "mortar_and_pestle", "pasta_maker", "ramekin", "skewer", "slow_cooker", "smoker", "wok", "blender", "candy_thermometer", "pressure_cooker")

cooking_techniques <- c("raw", "saute", "freeze_chill", "fry", "stir_fry", "simmer", "boil", "broil", "bake", "braise", "chill", "deep_fry", "steam", "rub", "roast", "poach", "pan_fry", "marinate", "grill_barbecue", "grill")

nutritional_values <- c("calories", "protein", "fat", "sodium")

repice_type <- c("aperitif", "appetizer", "backyard_bbq", "breakfast", "brunch", "dessert", "dinner", "lunch", "side", "snack")

diet_preferences <- c( "dairy_free", "fat_free", "kosher","kosher_for_passover", "low_carb", "low_fat", "low_sodium", "low_sugar", "low_no_sugar", "non_alcoholic", "no_sugar_added", "organic", "soy_free", "tree_nut_free", "vegan", "vegetarian")

#ingredients
#low level categories

vegetables_vec <- c("artichoke", "arugula", "asparagus", "butternut_squash", "bean", "beet", "bell_pepper", "bok_choy", "broccoli", "broccoli_rabe", "brussel_sprout", "cabbage", "capers", "carrot", "cauliflower", "celery", "chard", "chile_pepper", "collard_greens", "corn", "cucumber", "eggplant", "endive", "escarole", "fennel", "garlic", "ginger", "green_bean", "green_onion_scallion", "horseradish", "jerusalem_artichoke", "jicama", "kale", "leafy_green", "leek", "legume", "lentil", "lettuce", "lima_bean", "mushroom", "mustard_greens", "okra", "onion", "parsnip", "pea", "pickles", "poblano", "pumpkin", "radicchio", "radish", "root_vegetable", "rutabaga", "salad", "shallot", "soy", "spinach", "squash", "sugar_snap_pea", "tapioca", "tomatillo", "tomato", "turnip", "watercress", "yellow_squash", "yuca", "zucchini")

pork_meat_vec <- c("bacon", "ham", "pork", "pork_chop", "pork_rib", "pork_tenderloin", "prosciutto")

lamb_meat_vec <- c("ground_lamb", "lamb", "lamb_chop", "lamb_shank", "rack_of_lamb")

beef_meat_vec <- c("beef", "beef_rib", "beef_shank", "beef_tenderloin", "brisket", "ground_beef", "hamburger", "veal")

meat_with_wings_vec <- c("chicken", "duck", "goose", "poultry", "poultry_sausage", "quail", "turkey")

meat_various_vec <- c("meatball", "meatloaf", "rabbit", "sausage", "steak", "venison")

# stuff_in_the_water <- c("anchovy", "bass", "caviar", "clam", "cod", "crab", "fish", "halibut", "lobster", "mussel", "octopus", "oyster", "salmon", "sardine", "scallop", "seafood", "shellfish", "shrimp", "snapper", "squid", "swordfish", "tilapia", "trout", "tuna")
  
seafood_vec <- c("clam", "crab", "lobster", "mussel", "octopus", "oyster", "scallop", "shellfish", "shrimp", "squid")

fish_vec <- c("anchovy", "bass", "caviar", "cod", "halibut", "salmon", "sardine", "snapper", "swordfish", "tilapia", "trout", "tuna")
  

herbs_vec <- c("anise", "basil", "chive", "cilantro", "coriander", "dill", "lemongrass", "mint", "oregano", "parsley", "rosemary", "sage", "tarragon", "thyme")

nuts_vec <- c("almond", "cashew", "chestnut", "hazelnut", "macadamia_nut", "peanut", "pecan", "pine_nut", "pistachio", "tree_nut", "walnut")

cereals_vec <- c("barley", "bran", "bulgur", "grains", "granola", "oat", "quinoa", "rye", "whole_wheat")

carbs_vec <- c("brown_rice", "chickpea", "cornmeal", "couscous", "hominy_cornmeal_masa", "orzo", "pasta", "potato", "rice", "semolina", "sweet_potato_yam", "wild_rice")

fruits_vec <- c("apple", "apricot", "asian_pear", "avocado", "banana", "berry", "blackberry", "blueberry", "cantaloupe", "cherry", "citrus", "coconut", "cranberry", "currant", "dried_fruit", "fig", "grape", "grapefruit", "guava", "honeydew", "kiwi", "kumquat", "lemon", "lime", "lingonberry", "lychee", "mango", "melon", "nectarine", "olive", "orange", "papaya", "passion_fruit", "peach", "pear", "persimmon", "pineapple", "plantain", "plum", "pomegranate", "prune", "quince", "raisin", "raspberry", "rhubarb", "strawberry", "tamarind", "tangerine", "tropical_fruit", "watermelon")

drink_no_alcohol_vec <- c("apple_juice", "fruit_juice", "iced_tea", "lemon_juice", "lime_juice", "orange_juice", "pomegranate_juice", "tea")

dessert_vec <- c("biscuit", "brownie", "butterscotch_caramel", "cake", "candy", "chocolate", "cobbler_crumble", "cookie", "cookies", "cranberry_sauce", "crepe", "cupcake", "honey", "jam_or_jelly", "maple_syrup", "marshmallow", "muffin","pancake", "pastry", "pie", "smoothie", "sorbet", "souffle_meringue", "waffle")

cheeses_vec <- c("blue_cheese", "brie", "cheddar", "cottage_cheese", "cream_cheese", "feta", "fontina", "goat_cheese", "gouda", "monterey_jack", "mozzarella", "parmesan", "ricotta", "swiss_cheese")

dairy_vec <- c("butter", "buttermilk", "custard", "egg_nog", "ice_cream", "marscarpone", "milk_cream", "sour_cream", "yogurt")

spices_vec <- c("caraway", "cardamom", "chili", "cinnamon", "clove", "cumin", "curry", "hot_pepper", "jalapeno", "marinade", "nutmeg", "paprika", "pepper", "poppy", "saffron", "sesame", "sesame_oil", "soy_sauce", "vanilla", "wasabi")

#top level categories
general_categories <- c("vegetable", "meat", "fish", "seafood", "herb", "nut", "fruit", "drink", "cheese", "dairy", "spice")#using this to select the columns in ingredients_df and we could also use it later of for the for loop
  
all_meats <- c(beef_meat_vec, pork_meat_vec, lamb_meat_vec, meat_with_wings_vec, meat_various_vec)

all_fish_seafood <- c(fish_vec, seafood_vec)

all_ingredients <- c(vegetables_vec, all_meats, all_fish_seafood, herbs_vec, nuts_vec, cereals_vec, carbs_vec, fruits_vec, drink_no_alcohol_vec, dessert_vec, cheeses_vec, dairy_vec, spices_vec, "egg")

#stuff which isn't ingredients and that we need to sort
to_sort <- c("bread", "breadcrumbs", "brine", "burrito", "casserole_gratin", "coffee", "flat_bread", "hummus", "iced_coffee", "lasagna", "macaroni_and_cheese", "mayonnaise", "mustard", "noodle", "oatmeal", "omelet",  "peanut_butter", "pizza", "pot_pie", "potato_salad", "quiche", "rose", "salad_dressing", "salsa", "sandwich", "sauce", "seed", "soup_stew", "stew", "stock", "stuffing_dressing", "taco", "tart", "tofu", "tortillas", "vinegar", "frittata", "molasses", "sourdough", "fritter", "phyllo_puff_pastry_dough", "dip")
```


# TEMPORARY TO MAKE VARIABLE SELECTION EASIER

```{r}
#whole list of stuff to remove for now to be able to sort 
to_remove_temp <- c(us_states, us_cities, countries, alcohol, wtf, chef, interesting, season, celebrations, tools, cooking_techniques, nutritional_values, repice_type, diet_preferences, all_ingredients, to_sort)

#tried this with select but didn't work because some columns in the vector don't exist in the dataset
recipes_to_filter <- recipes[, !(colnames(recipes) %in% to_remove_temp)]

#creates a tibble with one column with all the colnames to be able to sort ingredients
names <- recipes_to_filter %>% colnames() %>% as_tibble()

# recipes %>% 
#   select(iceland)
  # filter(x_cakeweek==1)

#checked if some values in our categories weren't columns in recipes --> 19 of them weren't so I deleted them from the vectors
checking <- tibble(to_remove_temp[!to_remove_temp %in% colnames(recipes)])
checking


#checking only for ingredients because I feel like there is one in the ingredient vector which is not a column name
checking_ing <- tibble(all_ingredients[!all_ingredients %in% colnames(recipes)])
checking_ing

#no name comes out so the only logical conclusion is that there is a duplicated ingrediant name --> it was arugula
all_ingredients[duplicated(all_ingredients)]

```
