# Data Description and Cleaning

## Loading data
Here we load the dataset and do some cleaning and processing. We standardise the variables names and add an ID column to have a unique identifier for each recipe.
```{r Loading and early cleaning, code_folding = show}
#loading the data
recipes_raw <- read.csv(here("data/epi_r.csv"))

recipes <- recipes_raw%>% 
  clean_names() %>% 
  mutate(ID = 1:nrow(.)) %>%
  select(ID, everything())
```

## Example of the data
```{r}

selected_columns <- names(recipes)[recipes[1, ] == 1]
additional_variables <- c("title", "rating", "calories", "protein", "fat", "sodium")
selected_columns <- unique(c(additional_variables, selected_columns))

recipes[, selected_columns] %>% 
  filter(ID==1)
  

```
Let us focus on the first recipe as an example and take a look at the different variables of interest. We can observe the title as well as the numerical variables such as rating, calories, protein, fat and sodium. We will then see among the many categorical variables, those that are related to our recipe such as apple, bean, cookie, fruit, kid_friendly and so on.

## Data Description
```{r}
tibble(Variables = c("**ID**", "**title**", "**rating**", "**calories**", "**protein**", "**fat**", "**sodium**", "**674 other binary variables**"), Meaning = c("Unique ID", "Recipe name", "Rating of the recipe", "Calories contained in the recipe", "Protein contained in the recipe (grams)","Fat contained in the recipe (grams)", "Sodium contained in the recipe (milligrams)", "The rest of the data is made of many binary variables, incl. ingredients, types of recipes, US States, diet preferences, etc."))%>%
  kbl()%>%
  kable_styling(position = "center")
```

## Classifying variables into categories

Given the high amount of variables that we had (680), we decided that we needed to somewhat create categories to aggregate them and be able to use them more easily.

Specifically, here are the things we had to solve when creating categories:

- Merge "father_s_day" and "fathers_day" same for mother's day and for new year's and st patrick and valentines day, vermout and vermouth - DONE
- Check what leafy_green is and how many obs there is of it
- Same for "legume"
- Decide if we group beans together or leave them in vegetables - DONE
- What do we do with "meat", meatball and meatloaf. What about rabbit and for sausage or steak and venison - DONE
- Put nutmeg in spices or nuts? - DONE
- do we create a seeds category --> for example for "poppy" (put in spices for now) and for "seed" and sesame - DONE
- We should probably create a sauce category, and a "full_meal"
- where do we put tapioca and yuca - DONE
- do we put buttermilk in drinks or dairy?
- check if dorie_greenspa column exist or if it was just a typo without the N
- check how many observations with phyllo_puff_pastry_dough
- do we separate fish and seafood?

We have now solved all these issues and manually classified every variable of our dataset in specific categories shown below.

```{r Creating Variable Categories}
us_states <- c("alabama", "alaska", "arizona", "california", "colorado", "connecticut", "florida", "georgia", "hawaii", "idaho", "illinois", "indiana", "iowa", "kansas", "kentucky", "louisiana", "maine", "maryland", "massachusetts", "michigan", "minnesota", "mississippi", "missouri", "nebraska", "new_hampshire", "new_jersey", "new_mexico", "new_york", "north_carolina", "ohio", "oklahoma", "oregon", "pennsylvania", "rhode_island", "south_carolina", "tennessee", "texas", "utah", "vermont", "virginia", "washington", "west_virginia", "wisconsin")

us_cities <- c("aspen", "atlanta", "beverly_hills", "boston","brooklyn", "buffalo", "cambridge", "chicago", "columbus", "costa_mesa", "dallas", "denver", "healdsburg", "hollywood", "houston", "kansas_city", "lancaster", "las_vegas", "london", "los_angeles", "louisville", "miami", "minneapolis", "new_orleans", "pacific_palisades", "paris", "pasadena", "pittsburgh", "portland", "providence", "san_francisco", "santa_monica", "seattle", "st_louis", "washington_d_c", "westwood", "yonkers")

countries <- c("australia", "bulgaria", "canada", "chile", "cuba", "dominican_republic", "egypt", "england", "france", "germany", "guam", "haiti", "ireland", "israel", "italy", "jamaica", "japan", "mexico", "mezcal", "peru", "philippines", "spain", "switzerland")

alcohol <- c("alcoholic", "amaretto", "beer", "bitters", "bourbon", "brandy", "calvados", "campari", "chambord", "champagne", "chartreuse", "cocktail", "cognac_armagnac", "creme_de_cacao", "digestif", "eau_de_vie", "fortified_wine", "frangelico", "gin", "grand_marnier", "grappa", "kahlua", "kirsch", "liqueur", "long_beach", "margarita", "marsala", "martini", "midori", "pernod", "port", "punch", "red_wine", "rum", "sake", "sangria", "scotch", "sherry", "sparkling_wine", "spirit", "spritzer", "tequila", "triple_sec", "vermouth", "vodka", "whiskey", "white_wine", "wine")

wtf <- c("bon_appetit", "bon_app_tit", "condiment_spread", "cr_me_de_cacao", "epi_ushg", "flaming_hot_summer", "frankenrecipe", "harpercollins", "house_garden", "no_meat_no_problem", "parade", "sandwich_theory", "self", "shower", "tested_improved", "windsor", "weelicious", "snack_week", "tailgating", "quick_and_healthy", "picnic", "kitchen_olympics", "house_cocktail", "hors_d_oeuvre", "frozen_dessert", "freezer_food", "edible_gift", "cookbook_critic", "cook_like_a_diner", "condiment", "cocktail_party", "camping", "buffet", "x30_days_of_groceries", "x_cakeweek", "x_wasteless", "x22_minute_meals", "x3_ingredient_recipes")

chef <- c("anthony_bourdain", "dorie_greenspan", "emeril_lagasse", "nancy_silverton", "suzanne_goin")

interesting <- c("advance_prep_required", "entertaining", "epi_loves_the_microwave", "friendsgiving", "game", "gourmet", "healthy", "high_fiber", "hot_drink", "kid_friendly", "kidney_friendly", "low_cal", "low_cholesterol", "microwave", "no_cook", "one_pot_meal", "oscars", "paleo", "pescatarian", "poker_game_night", "potluck", "quick_easy", "cookbooks", "leftovers")

seasons_vec <- c("winter", "spring", "summer", "fall")

celebrations <- c("anniversary", "back_to_school", "bastille_day", "birthday", "christmas", "christmas_eve", "cinco_de_mayo", "date", "diwali", "easter", "engagement_party", "family_reunion", "father_s_day", "fourth_of_july", "graduation", "halloween", "hanukkah", "kentucky_derby", "kwanzaa", "labor_day", "lunar_new_year", "mardi_gras", "mother_s_day", "new_year_s_day", "new_year_s_eve", "oktoberfest", "party", "passover", "persian_new_year", "purim", "ramadan", "rosh_hashanah_yom_kippur", "shavuot", "st_patrick_s_day", "sukkot", "super_bowl", "thanksgiving", "valentine_s_day", "wedding")

drink_no_alcohol_vec <- c("apple_juice", "fruit_juice", "iced_tea", "lemon_juice", "lime_juice", "orange_juice", "pomegranate_juice", "tea")

tools <- c("coffee_grinder", "double_boiler", "food_processor", "ice_cream_machine", "juicer", "mandoline", "mixer", "mortar_and_pestle", "pasta_maker", "ramekin", "skewer", "slow_cooker", "smoker", "wok", "blender", "candy_thermometer", "pressure_cooker")

cooking_techniques <- c("raw", "saute", "freeze_chill", "fry", "stir_fry", "simmer", "boil", "broil", "bake", "braise", "chill", "deep_fry", "steam", "rub", "roast", "poach", "pan_fry", "marinate", "grill_barbecue", "grill")

nutritional_values <- c("calories", "protein", "fat", "sodium")

recipe_type_vec <- c("aperitif", "appetizer", "breakfast", "brunch", "dessert", "dinner", "lunch", "side", "snack")

diet_preferences_vec <- c( "dairy_free", "fat_free", "kosher","kosher_for_passover", "low_carb", "low_fat", "low_sodium", "low_sugar", "low_no_sugar", "non_alcoholic", "no_sugar_added", "organic", "peanut_free", "soy_free", "sugar_conscious", "tree_nut_free", "vegan", "vegetarian", "wheat_gluten_free")

### Ingredients
#low level categories

vegetables_vec <- c("artichoke", "arugula", "asparagus", "butternut_squash", "bean", "beet", "bell_pepper", "bok_choy", "broccoli", "broccoli_rabe", "brussel_sprout", "cabbage", "capers", "carrot", "cauliflower", "celery", "chard", "chile_pepper", "collard_greens", "corn", "cucumber", "eggplant", "endive", "escarole", "fennel", "garlic", "ginger", "green_bean", "green_onion_scallion", "horseradish", "jerusalem_artichoke", "jicama", "kale", "leafy_green", "leek", "legume", "lentil", "lettuce", "lima_bean", "mushroom", "mustard_greens", "okra", "onion", "parsnip", "pea", "pickles", "poblano", "pumpkin", "radicchio", "radish", "root_vegetable", "rutabaga", "salad", "shallot", "soy", "spinach", "squash", "sugar_snap_pea", "tapioca", "tomatillo", "tomato", "turnip", "watercress", "yellow_squash", "yuca", "zucchini")

pork_meat_vec <- c("bacon", "ham", "pork", "pork_chop", "pork_rib", "pork_tenderloin", "prosciutto")

lamb_meat_vec <- c("ground_lamb", "lamb", "lamb_chop", "lamb_shank", "rack_of_lamb")

beef_meat_vec <- c("beef", "beef_rib", "beef_shank", "beef_tenderloin", "brisket", "ground_beef", "hamburger", "veal")

meat_with_wings_vec <- c("chicken", "duck", "goose", "poultry", "poultry_sausage", "quail", "turkey")

meat_various_vec <- c("meatball", "meatloaf", "rabbit", "sausage", "steak", "venison")

# stuff_in_the_water <- c("anchovy", "bass", "caviar", "clam", "cod", "crab", "fish", "halibut", "lobster", "mussel", "octopus", "oyster", "salmon", "sardine", "scallop", "seafood", "shellfish", "shrimp", "snapper", "squid", "swordfish", "tilapia", "trout", "tuna")
  
seafood_vec <- c("clam", "crab", "lobster", "mussel", "octopus", "oyster", "scallop", "shellfish", "shrimp", "squid")

fish_vec <- c("anchovy", "bass", "caviar", "cod", "halibut", "salmon", "sardine", "snapper", "swordfish", "tilapia", "trout", "tuna")
  

herbs_vec <- c("anise", "basil", "chive", "cilantro", "coriander", "dill", "lemongrass", "mint", "oregano", "parsley", "rosemary", "sage", "tarragon", "thyme")

nuts_vec <- c("almond", "cashew", "chestnut", "hazelnut", "macadamia_nut", "peanut", "pecan", "pine_nut", "pistachio", "tree_nut", "walnut")

cereals_vec <- c("barley", "bran", "bulgur", "grains", "granola", "oat", "quinoa", "rye", "whole_wheat")

carbs_vec <- c("brown_rice", "chickpea", "cornmeal", "couscous", "hominy_cornmeal_masa", "orzo", "pasta", "potato", "rice", "semolina", "sweet_potato_yam", "wild_rice")

fruits_vec <- c("apple", "apricot", "asian_pear", "avocado", "banana", "berry", "blackberry", "blueberry", "cantaloupe", "cherry", "citrus", "coconut", "cranberry", "currant", "dried_fruit", "fig", "grape", "grapefruit", "guava", "honeydew", "kiwi", "kumquat", "lemon", "lime", "lingonberry", "lychee", "mango", "melon", "nectarine", "olive", "orange", "papaya", "passion_fruit", "peach", "pear", "persimmon", "pineapple", "plantain", "plum", "pomegranate", "prune", "quince", "raisin", "raspberry", "rhubarb", "strawberry", "tamarind", "tangerine", "tropical_fruit", "watermelon")

dessert_vec <- c("biscuit", "brownie", "butterscotch_caramel", "cake", "candy", "chocolate", "cobbler_crumble", "cookie", "cookies", "cranberry_sauce", "crepe", "cupcake", "honey", "jam_or_jelly", "maple_syrup", "marshmallow", "muffin","pancake", "pastry", "pie", "smoothie", "sorbet", "souffle_meringue", "waffle")

cheeses_vec <- c("blue_cheese", "brie", "cheddar", "cottage_cheese", "cream_cheese", "feta", "fontina", "goat_cheese", "gouda", "monterey_jack", "mozzarella", "parmesan", "ricotta", "swiss_cheese")

dairy_vec <- c("butter", "buttermilk", "custard", "egg_nog", "ice_cream", "marscarpone", "milk_cream", "sour_cream", "yogurt")

spices_vec <- c("caraway", "cardamom", "chili", "cinnamon", "clove", "cumin", "curry", "hot_pepper", "jalapeno", "marinade", "nutmeg", "paprika", "pepper", "poppy", "saffron", "sesame", "sesame_oil", "soy_sauce", "vanilla", "wasabi")

#top level categories
general_categories <- c("vegetable", "meat", "fish", "seafood", "herb", "nut", "fruit", "drink", "cheese", "dairy", "spice")#using this to select the columns in ingredients_df and we could also use it later of for the for loop
  
all_meats <- c(beef_meat_vec, pork_meat_vec, lamb_meat_vec, meat_with_wings_vec, meat_various_vec)

all_fish_seafood <- c(fish_vec, seafood_vec)

all_ingredients <- c(vegetables_vec, all_meats, all_fish_seafood, herbs_vec, nuts_vec, cereals_vec, carbs_vec, fruits_vec, drink_no_alcohol_vec, dessert_vec, cheeses_vec, dairy_vec, spices_vec, "egg")

#stuff which isn't ingredients and that we need to sort
to_sort <- c("backyard_bbq", "bread", "breadcrumbs", "brine", "burrito", "casserole_gratin", "coffee", "flat_bread", "hummus", "iced_coffee", "lasagna", "macaroni_and_cheese", "mayonnaise", "mustard", "noodle", "oatmeal", "omelet",  "peanut_butter", "pizza", "pot_pie", "potato_salad", "quiche", "rose", "salad_dressing", "salsa", "sandwich", "sauce", "seed", "soup_stew", "stew", "stock", "stuffing_dressing", "taco", "tart", "tofu", "tortillas", "vinegar", "frittata", "molasses", "sourdough", "fritter", "phyllo_puff_pastry_dough", "dip")
```


```{r TEMPORARY TO MAKE VARIABLE SELECTION EASIER, eval = FALSE}
#whole list of stuff to remove for now to be able to sort 
to_remove_temp <- c(us_states, us_cities, countries, alcohol, wtf, chef, interesting, season, celebrations, tools, cooking_techniques, nutritional_values, repice_type, diet_preferences, all_ingredients, to_sort)

#tried this with select but didn't work because some columns in the vector don't exist in the dataset
recipes_to_filter <- recipes[, !(colnames(recipes) %in% to_remove_temp)]

#creates a tibble with one column with all the colnames to be able to sort ingredients
names <- recipes_to_filter %>% colnames() %>% as_tibble()

# recipes %>% 
#   select(iceland)
  # filter(x_cakeweek==1)

#checked if some values in our categories weren't columns in recipes --> 19 of them weren't so I deleted them from the vectors
checking <- tibble(to_remove_temp[!to_remove_temp %in% colnames(recipes)])
checking


#checking only for ingredients because I feel like there is one in the ingredient vector which is not a column name
checking_ing <- tibble(all_ingredients[!all_ingredients %in% colnames(recipes)])
checking_ing

#no name comes out so the only logical conclusion is that there is a duplicated ingrediant name --> it was arugula
all_ingredients[duplicated(all_ingredients)]

```

# Data Understanding

## Structure and summary

```{r}

# Now let's see the structure of our data
recipes %>% 
  head(20) %>% 
  str() 

# We have only numerical variables, but in reality just 4 variables could be considered as such. More in particular, "rating", "calories", "protein", "fat" and "sodium" could be considered numerical. The other variables should be considered categorical since they allow only for 0 or 1 values.


# Let's have a different look at the data with the summary function.
recipes %>% 
  select(rating, calories, protein, fat, sodium) %>% 
  dfSummary(style = "grid")

# We can already see for instance that the majority of the values of the variable "rating" are 4.38 (40% of the total). Moreover, we observe that the variables "calories", "protein", "fat" and "sodium" have roughly 20% of missing values.

```

# Data Cleaning

## Analysis of NAs

```{r}
#we use the temp_df each time we want to create a temporary df for a single analysis and we know we won't reuse that dataframe later on
temp_df <- recipes %>% 
  select(title, all_of(nutritional_values))


na_obs <- which(rowSums(is.na(temp_df)) > 0)

# subset the original dataframe to only include rows with NA values
df_na <- temp_df[na_obs, ]

# print the result
#df_na


# count the number of NAs for each row
na_count <- rowSums(is.na(df_na))

# count the frequency of NA counts
freq_table_na <- table(na_count)

freq_na <- as.data.frame(freq_table_na) %>% 
  mutate(na_count = as.character(na_count))


freq_na %>% 
  ggplot(aes(x=na_count, y=Freq)) +
  geom_bar(stat="identity") +
  xlab("Number of NAs") +
  ylab("Frequency") +
  labs(title ="Number of NAs in nutritional values per recipe", subtitle = "NAs among variables calories, protein, fat, sodium") +
  coord_flip()


recipes <- recipes %>% 
  drop_na()

```
Among the recipes which have NAs, we notice that many of them have 4 NAs for all the 4 nutritional values, more precisely 4117  out of 4188 recipes. Without any other information available, making an imputation to retrieve such values would not make any sense. Interestingly, we do not observe 3 contemporary NAs for recipes.

We could try to make an imputation of the 29 recipes that have only 1 NA. The same operation on the 42 recipes with 2 NAs would not deliver accurate and satisfying results. However, we believe that is not worth to make imputation of such NA values. We should not forget that the nutritional values per recipe are estimated, then making an imputation would result in a sort of estimation of an estimation. To what extent could it be reliable? We decide to eliminate recipes with NAs for nutritional values. Nutritional values represent a crucial information for our analysis. 

Finally, we would still have 15864 recipes without NAs.


## Eliminate recipes with rating equal to zero

```{r}

rating_count <- table(recipes$rating) %>% 
  as.data.frame() %>% 
  rename(rating = Var1,
    frequency = Freq)

recipes <- recipes %>% 
  filter(rating != 0)


```

There are 1296 recipes which have rating equal to zero. Some of those might be unpopular, others might be too recent to have a rating. For the purpose of our analysis, we decide to eliminate these specific recipes.

We are left with 14568 observations after removing NAs and obs with a 0 rating value.

## Discard copies of recipes

We want to eliminate recipes that have multiple copies. Sometimes the recipes have the same title, but nutritional values are different. This indicates that there are various ways to prepare a specific recipe. We want to keep those recipes that have the same title, but have different nutritional values.

Let's check for instance Almond Butter Crisps, a recipe which can be found twice in the data set, with ID=1026 and ID=8908.
```{r}

# recipes %>% 
#   filter(ID == 1026)

unique_recipes <- distinct(recipes, title, rating, protein, sodium, fat, calories, .keep_all = TRUE)

# unique_recipes %>% 
#   filter(ID == 8908)

recipes <- unique_recipes

```

Now the data set is free from useless copies. We discarded 1288 copies in total. We lose a bit less if we remove the ones without specific ingredients first, meaning that some duplicate copies don't contain specific ingredients either.


## Removing recipes without specific ingredients listed
Here we are facing a challenge regarding the general ingredient categories. Indeed, when doing computations on the binary columns, there is no issue since, whether the recipe contains specific ingredients in a category, or only a 1 in the general category, then that information is captured in the corresponding binary column.

However, if we want to compute the total number of ingredients in each category that is present in recipes, then we are facing problems. To illustrate, let's assume that we have a recipe which contains 3 vegetables (specific columns in the vegetables_vec). In addition, for that recipe, the general column is also a 1 --> then by summing up, we get 4 ingredients when it should be 3.

Another problem is related to recipes for which the only column in a category (e.g., vegetables) that has a 1 is the general category (i.e., vegetable), and there isn't any specific ingredient listed within the vegetable category (in vegetables_vec) --> this can lead to issues when counting the number of specific ingredients per category.

In order to decide whether we want to analyse with or without general categories, let's see how many observations would remain if we remove all the obs for which we have a general category at 1, and all specific ingredients in that category is set to 0.

```{r Removing recipes without ingredients specified}
#this filters out the observations which have 1 for general category and 0s for every ingredient in that category
recipes <- recipes %>%
  filter(!(if_all(all_of(vegetables_vec), ~.x == 0) & vegetable == 1)) %>% 
  filter(!(if_all(all_of(all_meats), ~.x == 0) & meat == 1)) %>% 
  filter(!(if_all(all_of(fish_vec), ~.x == 0) & fish == 1)) %>% 
  filter(!(if_all(all_of(seafood_vec), ~.x == 0) & seafood == 1)) %>% 
  filter(!(if_all(all_of(herbs_vec), ~.x == 0) & herb == 1)) %>% 
  filter(!(if_all(all_of(nuts_vec), ~.x == 0) & nut == 1)) %>% 
  filter(!(if_all(all_of(fruits_vec), ~.x == 0) & fruit == 1)) %>% 
  filter(!(if_all(all_of(drink_no_alcohol_vec), ~.x == 0) & drink == 1)) %>% 
  filter(!(if_all(all_of(cheeses_vec), ~.x == 0) & cheese == 1)) %>% 
  filter(!(if_all(all_of(dairy_vec), ~.x == 0) & dairy == 1)) %>% 
  filter(!(if_all(all_of(spices_vec), ~.x == 0) & spice == 1))
```
We are left with 10321 obs after removing all recipes which have no specific ingredient in at least one category, while that category general variable is at 1.
