---
title: "Data"
output: html_document
date: "2023-04-20"
---

```{r echo=FALSE, message=FALSE}
source(here::here("scripts/setup.R"))
```
# Data Description and Cleaning

## Loading data
Here we load the dataset and do some cleaning and processing. We standardise the variables names and add an ID column to have a unique identifier for each recipe.
```{r Loading and early cleaning, code_folding = show}
#loading the data
recipes_raw <- read.csv(here("data/epi_r.csv"))

recipes <- recipes_raw%>% 
  clean_names() %>% 
  mutate(ID = 1:nrow(.)) %>%
  select(ID, everything())
```

## Data Description
```{r}
tibble(Variables = c("**ID**", "**title**", "**rating**", "**calories**", "**protein**", "**fat**", "**sodium**", "**674 other binary variables**"), Meaning = c("Unique ID", "Recipe name", "Rating of the recipe", "Calories contained in the recipe", "Protein contained in the recipe (grams)","Fat contained in the recipe (grams)", "Sodium contained in the recipe (milligrams)", "The rest of the data is made of many binary variables, incl. ingredients, types of recipes, US States, diet preferences, etc."))%>%
  kbl()%>%
  kable_styling(position = "center")
```

## Classifying variables into categories

Given the high amount of variables that we had (680), we decided that we needed to somewhat create categories to aggregate them and be able to use them more easily.

Specifically, here are the things we had to solve when creating categories:

- Merge "father_s_day" and "fathers_day" same for mother's day and for new year's and st patrick and valentines day, vermout and vermouth - DONE
- Check what leafy_green is and how many obs there is of it
- Same for "legume"
- Decide if we group beans together or leave them in vegetables - DONE
- What do we do with "meat", meatball and meatloaf. What about rabbit and for sausage or steak and venison - DONE
- Put nutmeg in spices or nuts? - DONE
- do we create a seeds category --> for example for "poppy" (put in spices for now) and for "seed" and sesame - DONE
- We should probably create a sauce category, and a "full_meal"
- where do we put tapioca and yuca - DONE
- do we put buttermilk in drinks or dairy?
- check if dorie_greenspa column exist or if it was just a typo without the N
- check how many observations with phyllo_puff_pastry_dough
- do we separate fish and seafood?

We have now solved all these issues and manually classified every variable of our dataset in specific categories shown below.

```{r Creating Variable Categories}
us_states <- c("alabama", "alaska", "arizona", "california", "colorado", "connecticut", "florida", "georgia", "hawaii", "idaho", "illinois", "indiana", "iowa", "kansas", "kentucky", "louisiana", "maine", "maryland", "massachusetts", "michigan", "minnesota", "mississippi", "missouri", "nebraska", "new_hampshire", "new_jersey", "new_mexico", "new_york", "north_carolina", "ohio", "oklahoma", "oregon", "pennsylvania", "rhode_island", "south_carolina", "tennessee", "texas", "utah", "vermont", "virginia", "washington", "west_virginia", "wisconsin")

us_cities <- c("aspen", "atlanta", "beverly_hills", "boston","brooklyn", "buffalo", "cambridge", "chicago", "columbus", "costa_mesa", "dallas", "denver", "healdsburg", "hollywood", "houston", "kansas_city", "lancaster", "las_vegas", "london", "los_angeles", "louisville", "miami", "minneapolis", "new_orleans", "pacific_palisades", "paris", "pasadena", "pittsburgh", "portland", "providence", "san_francisco", "santa_monica", "seattle", "st_louis", "washington_d_c", "westwood", "yonkers")

countries <- c("australia", "bulgaria", "canada", "chile", "cuba", "dominican_republic", "egypt", "england", "france", "germany", "guam", "haiti", "ireland", "israel", "italy", "jamaica", "japan", "mexico", "mezcal", "peru", "philippines", "spain", "switzerland")

alcohol <- c("alcoholic", "amaretto", "beer", "bitters", "bourbon", "brandy", "calvados", "campari", "chambord", "champagne", "chartreuse", "cocktail", "cognac_armagnac", "creme_de_cacao", "digestif", "eau_de_vie", "fortified_wine", "frangelico", "gin", "grand_marnier", "grappa", "kahlua", "kirsch", "liqueur", "long_beach", "margarita", "marsala", "martini", "midori", "pernod", "port", "punch", "red_wine", "rum", "sake", "sangria", "scotch", "sherry", "sparkling_wine", "spirit", "spritzer", "tequila", "triple_sec", "vermouth", "vodka", "whiskey", "white_wine", "wine")

wtf <- c("bon_appetit", "bon_app_tit", "condiment_spread", "cr_me_de_cacao", "epi_ushg", "flaming_hot_summer", "frankenrecipe", "harpercollins", "house_garden", "no_meat_no_problem", "parade", "sandwich_theory", "self", "shower", "tested_improved", "windsor", "weelicious", "snack_week", "tailgating", "quick_and_healthy", "picnic", "kitchen_olympics", "house_cocktail", "hors_d_oeuvre", "frozen_dessert", "freezer_food", "edible_gift", "cookbook_critic", "cook_like_a_diner", "condiment", "cocktail_party", "camping", "buffet", "x30_days_of_groceries", "x_cakeweek", "x_wasteless", "x22_minute_meals", "x3_ingredient_recipes")

chef <- c("anthony_bourdain", "dorie_greenspan", "emeril_lagasse", "nancy_silverton", "suzanne_goin")

interesting <- c("advance_prep_required", "entertaining", "epi_loves_the_microwave", "friendsgiving", "game", "gourmet", "healthy", "high_fiber", "hot_drink", "kid_friendly", "kidney_friendly", "low_cal", "low_cholesterol", "microwave", "no_cook", "one_pot_meal", "oscars", "paleo", "pescatarian", "poker_game_night", "potluck", "quick_easy", "cookbooks", "leftovers")

seasons_vec <- c("winter", "spring", "summer", "fall")

celebrations <- c("anniversary", "back_to_school", "bastille_day", "birthday", "christmas", "christmas_eve", "cinco_de_mayo", "date", "diwali", "easter", "engagement_party", "family_reunion", "father_s_day", "fourth_of_july", "graduation", "halloween", "hanukkah", "kentucky_derby", "kwanzaa", "labor_day", "lunar_new_year", "mardi_gras", "mother_s_day", "new_year_s_day", "new_year_s_eve", "oktoberfest", "party", "passover", "persian_new_year", "purim", "ramadan", "rosh_hashanah_yom_kippur", "shavuot", "st_patrick_s_day", "sukkot", "super_bowl", "thanksgiving", "valentine_s_day", "wedding")

drink_no_alcohol_vec <- c("apple_juice", "fruit_juice", "iced_tea", "lemon_juice", "lime_juice", "orange_juice", "pomegranate_juice", "tea")

tools <- c("coffee_grinder", "double_boiler", "food_processor", "ice_cream_machine", "juicer", "mandoline", "mixer", "mortar_and_pestle", "pasta_maker", "ramekin", "skewer", "slow_cooker", "smoker", "wok", "blender", "candy_thermometer", "pressure_cooker")

cooking_techniques <- c("raw", "saute", "freeze_chill", "fry", "stir_fry", "simmer", "boil", "broil", "bake", "braise", "chill", "deep_fry", "steam", "rub", "roast", "poach", "pan_fry", "marinate", "grill_barbecue", "grill")

nutritional_values <- c("calories", "protein", "fat", "sodium")

recipe_type_vec <- c("aperitif", "appetizer", "breakfast", "brunch", "dessert", "dinner", "lunch", "side", "snack")

diet_preferences_vec <- c( "dairy_free", "fat_free", "kosher","kosher_for_passover", "low_carb", "low_fat", "low_sodium", "low_sugar", "low_no_sugar", "non_alcoholic", "no_sugar_added", "organic", "peanut_free", "soy_free", "sugar_conscious", "tree_nut_free", "vegan", "vegetarian", "wheat_gluten_free")

### Ingredients
#low level categories

vegetables_vec <- c("artichoke", "arugula", "asparagus", "butternut_squash", "bean", "beet", "bell_pepper", "bok_choy", "broccoli", "broccoli_rabe", "brussel_sprout", "cabbage", "capers", "carrot", "cauliflower", "celery", "chard", "chile_pepper", "collard_greens", "corn", "cucumber", "eggplant", "endive", "escarole", "fennel", "garlic", "ginger", "green_bean", "green_onion_scallion", "horseradish", "jerusalem_artichoke", "jicama", "kale", "leafy_green", "leek", "legume", "lentil", "lettuce", "lima_bean", "mushroom", "mustard_greens", "okra", "onion", "parsnip", "pea", "pickles", "poblano", "pumpkin", "radicchio", "radish", "root_vegetable", "rutabaga", "salad", "shallot", "soy", "spinach", "squash", "sugar_snap_pea", "tapioca", "tomatillo", "tomato", "turnip", "watercress", "yellow_squash", "yuca", "zucchini")

pork_meat_vec <- c("bacon", "ham", "pork", "pork_chop", "pork_rib", "pork_tenderloin", "prosciutto")

lamb_meat_vec <- c("ground_lamb", "lamb", "lamb_chop", "lamb_shank", "rack_of_lamb")

beef_meat_vec <- c("beef", "beef_rib", "beef_shank", "beef_tenderloin", "brisket", "ground_beef", "hamburger", "veal")

meat_with_wings_vec <- c("chicken", "duck", "goose", "poultry", "poultry_sausage", "quail", "turkey")

meat_various_vec <- c("meatball", "meatloaf", "rabbit", "sausage", "steak", "venison")

# stuff_in_the_water <- c("anchovy", "bass", "caviar", "clam", "cod", "crab", "fish", "halibut", "lobster", "mussel", "octopus", "oyster", "salmon", "sardine", "scallop", "seafood", "shellfish", "shrimp", "snapper", "squid", "swordfish", "tilapia", "trout", "tuna")
  
seafood_vec <- c("clam", "crab", "lobster", "mussel", "octopus", "oyster", "scallop", "shellfish", "shrimp", "squid")

fish_vec <- c("anchovy", "bass", "caviar", "cod", "halibut", "salmon", "sardine", "snapper", "swordfish", "tilapia", "trout", "tuna")
  

herbs_vec <- c("anise", "basil", "chive", "cilantro", "coriander", "dill", "lemongrass", "mint", "oregano", "parsley", "rosemary", "sage", "tarragon", "thyme")

nuts_vec <- c("almond", "cashew", "chestnut", "hazelnut", "macadamia_nut", "peanut", "pecan", "pine_nut", "pistachio", "tree_nut", "walnut")

cereals_vec <- c("barley", "bran", "bulgur", "grains", "granola", "oat", "quinoa", "rye", "whole_wheat")

carbs_vec <- c("brown_rice", "chickpea", "cornmeal", "couscous", "hominy_cornmeal_masa", "orzo", "pasta", "potato", "rice", "semolina", "sweet_potato_yam", "wild_rice")

fruits_vec <- c("apple", "apricot", "asian_pear", "avocado", "banana", "berry", "blackberry", "blueberry", "cantaloupe", "cherry", "citrus", "coconut", "cranberry", "currant", "dried_fruit", "fig", "grape", "grapefruit", "guava", "honeydew", "kiwi", "kumquat", "lemon", "lime", "lingonberry", "lychee", "mango", "melon", "nectarine", "olive", "orange", "papaya", "passion_fruit", "peach", "pear", "persimmon", "pineapple", "plantain", "plum", "pomegranate", "prune", "quince", "raisin", "raspberry", "rhubarb", "strawberry", "tamarind", "tangerine", "tropical_fruit", "watermelon")

dessert_vec <- c("biscuit", "brownie", "butterscotch_caramel", "cake", "candy", "chocolate", "cobbler_crumble", "cookie", "cookies", "cranberry_sauce", "crepe", "cupcake", "honey", "jam_or_jelly", "maple_syrup", "marshmallow", "muffin","pancake", "pastry", "pie", "smoothie", "sorbet", "souffle_meringue", "waffle")

cheeses_vec <- c("blue_cheese", "brie", "cheddar", "cottage_cheese", "cream_cheese", "feta", "fontina", "goat_cheese", "gouda", "monterey_jack", "mozzarella", "parmesan", "ricotta", "swiss_cheese")

dairy_vec <- c("butter", "buttermilk", "custard", "egg_nog", "ice_cream", "marscarpone", "milk_cream", "sour_cream", "yogurt")

spices_vec <- c("caraway", "cardamom", "chili", "cinnamon", "clove", "cumin", "curry", "hot_pepper", "jalapeno", "marinade", "nutmeg", "paprika", "pepper", "poppy", "saffron", "sesame", "sesame_oil", "soy_sauce", "vanilla", "wasabi")

#top level categories
general_categories <- c("vegetable", "meat", "fish", "seafood", "herb", "nut", "fruit", "drink", "cheese", "dairy", "spice")#using this to select the columns in ingredients_df and we could also use it later of for the for loop
  
all_meats <- c(beef_meat_vec, pork_meat_vec, lamb_meat_vec, meat_with_wings_vec, meat_various_vec)

all_fish_seafood <- c(fish_vec, seafood_vec)

all_ingredients <- c(vegetables_vec, all_meats, all_fish_seafood, herbs_vec, nuts_vec, cereals_vec, carbs_vec, fruits_vec, drink_no_alcohol_vec, dessert_vec, cheeses_vec, dairy_vec, spices_vec, "egg")

#stuff which isn't ingredients and that we need to sort
to_sort <- c("backyard_bbq", "bread", "breadcrumbs", "brine", "burrito", "casserole_gratin", "coffee", "flat_bread", "hummus", "iced_coffee", "lasagna", "macaroni_and_cheese", "mayonnaise", "mustard", "noodle", "oatmeal", "omelet",  "peanut_butter", "pizza", "pot_pie", "potato_salad", "quiche", "rose", "salad_dressing", "salsa", "sandwich", "sauce", "seed", "soup_stew", "stew", "stock", "stuffing_dressing", "taco", "tart", "tofu", "tortillas", "vinegar", "frittata", "molasses", "sourdough", "fritter", "phyllo_puff_pastry_dough", "dip")
```


```{r TEMPORARY TO MAKE VARIABLE SELECTION EASIER, eval = FALSE}
#whole list of stuff to remove for now to be able to sort 
to_remove_temp <- c(us_states, us_cities, countries, alcohol, wtf, chef, interesting, season, celebrations, tools, cooking_techniques, nutritional_values, repice_type, diet_preferences, all_ingredients, to_sort)

#tried this with select but didn't work because some columns in the vector don't exist in the dataset
recipes_to_filter <- recipes[, !(colnames(recipes) %in% to_remove_temp)]

#creates a tibble with one column with all the colnames to be able to sort ingredients
names <- recipes_to_filter %>% colnames() %>% as_tibble()

# recipes %>% 
#   select(iceland)
  # filter(x_cakeweek==1)

#checked if some values in our categories weren't columns in recipes --> 19 of them weren't so I deleted them from the vectors
checking <- tibble(to_remove_temp[!to_remove_temp %in% colnames(recipes)])
checking


#checking only for ingredients because I feel like there is one in the ingredient vector which is not a column name
checking_ing <- tibble(all_ingredients[!all_ingredients %in% colnames(recipes)])
checking_ing

#no name comes out so the only logical conclusion is that there is a duplicated ingrediant name --> it was arugula
all_ingredients[duplicated(all_ingredients)]

```
