# Data Cleaning

## Analysis of NAs

```{r}

recipes_nutrition <- recipes %>% 
  select(title, calories, protein, fat, sodium)


na_obs <- which(rowSums(is.na(recipes_nutrition)) > 0)

# subset the original dataframe to only include rows with NA values
df_na <- recipes_nutrition[na_obs, ]

# print the result
#df_na


# count the number of NAs for each row
na_count <- rowSums(is.na(df_na))

# count the frequency of NA counts
freq_table_na <- table(na_count)

freq_na <- as.data.frame(freq_table_na) %>% 
  mutate(na_count = as.character(na_count))


freq_na %>% 
  ggplot(aes(x=na_count, y=Freq)) +
  geom_bar(stat="identity") +
  xlab("Number of NAs") +
  ylab("Frequency") +
  ggtitle("Number of NAs in nutritional values per recipe") +
  coord_flip()
  
# Among the recipes which have NAs, we notice that many of them have 4 NAs for all the 4 nutritional values, more precisely 4117  out of 4188 recipes. Without any other information available, making an imputation to retrieve such values would not make any sense.

# We could try to make an imputation of the 29 recipes that have only 1 NA. The same operation on the 42 recipes with 2 NAs would not deliver accurate and satisfying results. However, we believe that is not worth to make imputation of such NA values. We should not forget that the nutritional values per recipe are estimated, then making an imputation would result in a sort of estimation of an estimation. To what extent could it be reliable? We decide to eliminate recipes with NA values. Finally, we would still have 15864 recipes without NAs for our analysis.

recipes <- recipes %>% 
  drop_na()

#15864 obs after droping NAs
```


## Eliminate recipes with rating equal to zero

```{r}

rating_count <- table(recipes$rating)

rating_count <- as.data.frame(rating_count) %>% 
  rename(rating = Var1,
    frequency = Freq)

# There are 1293 recipes which have rating equal to zero. Some of those might be unpopular, others might be too recent to have a rating. For the purpose of our analysis, we decide to eliminate these specific recipes.

recipes <- recipes %>% 
  filter(rating != 0)

# We are left with 14568 observations after removing NAs and obs with a 0 rating value

```

## Discard copies of recipes
```{r}

# We want to eliminate recipes that have multiple copies. Sometimes the recipes have the same title, but nutritional values are different. This indicates that there are various ways to prepare a specific recipe. We want to keep those recipes that have the same title, but have different nutritional values.

# Let's check for instance Almond Butter Crisps, a recipe which can be found twice in the data set, with ID=1026 and ID=8908.

# recipes %>% 
#   filter(ID == 1026)

unique_recipes <- distinct(recipes, title, rating, protein, sodium, fat, calories, .keep_all = TRUE)

# unique_recipes %>% 
#   filter(ID == 8908)

# Now the data set is free from useless copies. We discarded 1288 copies in total. #we lose a bit less if we remove the ones without specific ingredients first, meaning that some duplicate copies don't contain specific ingredients either

recipes <- unique_recipes

```


## Removing recipes without specific ingredients listed
Here we are facing a challenge regarding the general ingredient categories. Indeed, when doing computations on the binary columns, there is no issue since, whether the recipe contains specific ingredients in a category, or only a 1 in the general category, then that information is captured in the corresponding binary column.

However, if we want to compute the total number of ingredients in each category that is present in recipes, then we are facing problems. To illustrate, let's assume that we have a recipe which contains 3 vegetables (specific columns in the vegetables_vec). In addition, for that recipe, the general column is also a 1 --> then by summing up, we get 4 ingredients when it should be 3.

Another problem is related to recipes for which the only column in a category (e.g., vegetables) that has a 1 is the general category (i.e., vegetable), and there isn't any specific ingredient listed within the vegetable category (in vegetables_vec) --> this can lead to issues when counting the number of specific ingredients per category.

In order to decide whether we want to analyse with or without general categories, let's see how many observations would remain if we remove all the obs for which we have a general category at 1, and all specific ingredients in that category is set to 0.

```{r Removing recipes without ingredients specified}
#this filters out the observations which have 1 for general category and 0s for every ingredient in that category
recipes <- recipes %>%
  filter(!(if_all(all_of(vegetables_vec), ~.x == 0) & vegetable == 1)) %>% 
  filter(!(if_all(all_of(all_meats), ~.x == 0) & meat == 1)) %>% 
  filter(!(if_all(all_of(fish_vec), ~.x == 0) & fish == 1)) %>% 
  filter(!(if_all(all_of(seafood_vec), ~.x == 0) & seafood == 1)) %>% 
  filter(!(if_all(all_of(herbs_vec), ~.x == 0) & herb == 1)) %>% 
  filter(!(if_all(all_of(nuts_vec), ~.x == 0) & nut == 1)) %>% 
  filter(!(if_all(all_of(fruits_vec), ~.x == 0) & fruit == 1)) %>% 
  filter(!(if_all(all_of(drink_no_alcohol_vec), ~.x == 0) & drink == 1)) %>% 
  filter(!(if_all(all_of(cheeses_vec), ~.x == 0) & cheese == 1)) %>% 
  filter(!(if_all(all_of(dairy_vec), ~.x == 0) & dairy == 1)) %>% 
  filter(!(if_all(all_of(spices_vec), ~.x == 0) & spice == 1))
#we are left with 11382 obs after removing all recipes which have no specific ingredient in at least one category, while that category general variable is at 1

### without duplicated recipes we are left at 10321 obs
```

# EDA

## Nutrution EDA

### Structure and summary

```{r}

# Now let's see the structure of our data
recipes %>% 
  head(20) %>% 
  str() 

# We have only numerical variables, but in reality just 4 variables could be considered as such. More in particular, "rating", "calories", "protein", "fat" and "sodium" could be considered numerical. The other variables should be considered categorical since they allow only for 0 or 1 values.


# Let's have a different look at the data with the summary function.
recipes %>% 
  select(rating, calories, protein, fat, sodium) %>% 
  dfSummary(style = "grid")

# We can already see for instance that the majority of the values of the variable "rating" are 4.38 (40% of the total). Moreover, we observe that the variables "calories", "protein", "fat" and "sodium" have roughly 20% of missing values.

```

### Visual exploration - Univariate Analysis

#### Rating Barplot

```{r}

recipes %>% 
  ggplot(aes(x=as.factor(rating), fill=as.factor(rating) )) +  
  geom_bar( ) +
  scale_fill_manual(values = c("red4", "red3", "orangered", "orange", "gold", "greenyellow", "green3", "green4") ) +
  theme(legend.position="none") +
  scale_y_continuous(breaks=seq(0,10000,1000)) +
  labs(x = "Rating", y = "Count", 
       title = "Overview of recipes' ratings")

# As we can see, most of the ratings have value above 3.75, more in particular we notice that most of the recipes are rates with 4.375.
# A zero rating indicates that the recipe has not been evaluated yet.

```


#### Calories - Boxplot and Histogram

```{r}

recipes_boxplot <-
  recipes %>% 
  pivot_longer(cols = c(calories, protein, fat, sodium),
               names_to = "nutrition",
               values_to = "n_value")

recipes_calories <- recipes_boxplot %>% 
  filter(nutrition == "calories")

recipes_calories %>%
  ggplot(aes(x=nutrition, y=n_value, fill=nutrition)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
    theme_light() +
    theme(legend.position="none",
          plot.title = element_text(size=11)) +
  ggtitle("Boxplot of calories nutritional value") +
  xlab("") +
  ylab("Value")
  

# We notice that there are recipes with more than 30'000'000 calories which are clearly outliers. We must then discard those values in order to continue with a meaningful analysis. Otherwise from a visual point of view we could not extract any relevant information. If we take out these recipes, we still have roughly 30 recipes which have more than 7000 calories. For the purpose of this precise visualisation, we want to exclude them to have a better overview of the boxplot, but we don't consider those as outliers.

recipes_calories %>% 
  select(title, nutrition, n_value) %>% 
  arrange(desc(n_value)) 

recipes_calories <- recipes_calories %>% 
  filter(n_value <= 7000)


# Calories boxplot

recipes_calories %>%
  ggplot(aes(x=nutrition, y=n_value, fill=nutrition)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
    theme_light() +
    theme(legend.position="none",
          plot.title = element_text(size=11)) +
  scale_y_continuous(breaks=seq(0,7000,500)) +
  ggtitle("Boxplot of calories nutritional value") +
  xlab("") +
  ylab("Calories") 


# Calories histogram

recipes_calories %>% 
  ggplot(aes(x=n_value)) +
  geom_histogram(binwidth=50, fill="red3", color="red3", alpha=0.9) +
  theme(plot.title = element_text(size=15)) +
  scale_x_continuous(breaks=seq(0,10000,1000)) +
  scale_y_continuous(breaks=seq(0,1750,250)) +
  ggtitle("Distribution of calories across all recipes") +
  xlab("Calories") +
  ylab("Count") 

```

We can observe that most of the recipes have between 200 and 600 calories. By checking with the histogram the distribution of calories, we observe that indeed most of the recipes have less than 1000 calories.


#### Protein - Boxplot and Histogram

```{r}

recipes_protein <- recipes_boxplot %>% 
  filter(nutrition == "protein")

recipes_protein %>%
  ggplot(aes(x=nutrition, y=n_value, fill=nutrition)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
    theme_light() +
    theme(legend.position="none",
          plot.title = element_text(size=11)) +
  ggtitle("Boxplot of protein nutritional value") +
  xlab("") +
  ylab("Value")
  

# We notice that there are recipes with more than 50'000 grams of protein which are clearly outliers. We must then discard those values in order to continue with a meaningful analysis. Otherwise from a visual point of view we could not extract any relevant information. By checking on the epicurious website recipes  with protein values above 1000, we also verified that the amount of proteins was not justified. We came to that conclusion by evaluating the average values of protein per 100grams of each ingredient in the specific recipe.

recipes_protein %>% 
  select(title, nutrition, n_value) %>% 
  arrange(desc(n_value)) 

recipes_protein <- recipes_protein %>% 
  filter(n_value <= 1000)


# Proteins boxplot

recipes_protein %>%
  ggplot( aes(x=nutrition, y=n_value, fill=nutrition)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
    theme_light() +
    theme(legend.position="none",
          plot.title = element_text(size=11)) +
  scale_y_continuous(breaks=seq(0,7000,25)) +
  ggtitle("Boxplot of protein nutritional value") +
  xlab("") +
  ylab("Proteins") 


# Proteins histogram

recipes_protein %>% 
  ggplot(aes(x=n_value)) +
  geom_histogram(binwidth=7, fill="red3", color="red3", alpha=0.9) +
  theme(plot.title = element_text(size=15)) +
  scale_x_continuous(breaks=seq(0,1000,25)) +
  scale_y_continuous(breaks=seq(0,7000,250)) +
  ggtitle("Distribution of proteins across all recipes") +
  xlab("Proteins") +
  ylab("Count") 

```

From the boxplot, we observe that most recipes have less than 30 grams of proteins. By plotting the histogram, we verify that this information is correct. We could even extend the range to 100 proteins per recipe. We assume that recipes with values above this threshold contain ingredients like meat, tuna, salmon or shrimps.


#### Sodium - Boxplot and Histogram

```{r}

recipes_sodium <- recipes_boxplot %>% 
  filter(nutrition == "sodium")

recipes_sodium %>%
  ggplot(aes(x=nutrition, y=n_value, fill=nutrition)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
    theme_light() +
    theme(legend.position="none",
          plot.title = element_text(size=11)) +
  ggtitle("Boxplot of sodium nutritional value") +
  xlab("") +
  ylab("Value")
  

# We notice that there are recipes with more than 100'000 milligrams of sodium which are clearly outliers. We must then discard those values in order to continue with a meaningful analysis. By conducting further research, we realize that sodium values above 30'000 are highly suspicious.

recipes_sodium %>% 
  select(title, nutrition, n_value) %>% 
  arrange(desc(n_value)) 

recipes_sodium <- recipes_sodium %>% 
  filter(n_value <= 30000)


# Sodium boxplot

recipes_sodium %>%
  ggplot(aes(x=nutrition, y=n_value, fill=nutrition)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
    theme_light() +
    theme(legend.position="none",
          plot.title = element_text(size=11)) +
  scale_y_continuous(breaks=seq(0,30000,500)) +
  ggtitle("Boxplot of sodium nutritional value") +
  xlab("") +
  ylab("Sodium") 


# Sodium histogram

recipes_sodium %>% 
  ggplot(aes(x=n_value)) +
  geom_histogram(binwidth=50, fill="red3", color="red3", alpha=0.9) +
  theme(plot.title = element_text(size=15)) +
  scale_x_continuous(breaks=seq(0,30000,1000)) +
  scale_y_continuous(breaks=seq(0,1750,250)) +
  ggtitle("Distribution of sodium across all recipes") +
  xlab("Sodium") +
  ylab("Count") 

```

From the boxplot we observe that most recipes have sodium values below 750 milligrams. The histogram informs us that most of recipes have indeed less than 750 milligrams of sodium, even though we cannot exclude the presence of a good amount of recipes with sodium between 750 and 2000 milligrams.


#### Fat - Boxplot and Histogram

```{r}

recipes_fat <- recipes_boxplot %>% 
  filter(nutrition == "fat")

recipes_fat %>%
  ggplot(aes(x=nutrition, y=n_value, fill=nutrition)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
    theme_light() +
    theme(legend.position="none",
          plot.title = element_text(size=11)) +
  ggtitle("Boxplot of fat nutritional value") +
  xlab("") +
  ylab("Value")
  

# We notice that there are recipes with more than 44'000 grams of fat which are clearly outliers. We must then discard those values in order to continue with a meaningful analysis. By checking on the epicurious website recipes  with fat values above 1000, we also verified that the amount of proteins was not justified. We came to that conclusion by evaluating the average values of protein per 100grams of each ingredient in the specific recipe.

recipes_fat %>% 
  select(title, nutrition, n_value) %>% 
  arrange(desc(n_value)) 

recipes_fat <- recipes_fat %>% 
  filter(n_value <= 40000)


# Fat boxplot

recipes_fat %>%
  ggplot( aes(x=nutrition, y=n_value, fill=nutrition)) +
    geom_boxplot() +
    scale_fill_viridis(discrete = TRUE, alpha=0.6, option="A") +
    theme_light() +
    theme(legend.position="none",
          plot.title = element_text(size=11)) +
  scale_y_continuous(breaks=seq(0,3000,100)) +
  ggtitle("Boxplot of fat nutritional value") +
  xlab("") +
  ylab("Fat") 


# Fat histogram

recipes_fat %>% 
  ggplot(aes(x=n_value)) +
  geom_histogram(binwidth=7, fill="red3", color="red3", alpha=0.9) +
  theme(plot.title = element_text(size=15)) +
  ggtitle("Distribution of fat across all recipes") +
  scale_x_continuous(breaks=seq(0,3000,100)) +
  scale_y_continuous(breaks=seq(0,7000,250)) +
  xlab("Fat") +
  ylab("Count") 

```

It is hard to interpret the boxplot. There are certain recipes which could have potentially more than 1000 or even 2000 grams of fat because of the high quantity of servings and the use of ingredients such as lamb, duck and bacon. We must then analyse the histogram to have a better overview and we notice that most recipes have fat values below 100 grams.



### Visual exploration - Multivariate Analysis

#### Scatterplots of Rating
```{r}
  
recipes <- recipes %>% 
  filter(calories <= 7000, protein <= 1000, sodium <= 30000, fat <= 40000)


# Scatterplot of Rating-Calories
recipes %>% 
  ggplot(aes(x=calories, y=rating)) +
  geom_point() +
  ggtitle("Scatterplot of rating against calories") +
  xlab("Calories") +
  ylab("Rating") 

# We can observe that the recipes with more than 2000 calories tend to have a higher rating. For instance, few recipes with less than a 3 star rating have more than 2000 calories.


# Scatterplot of Rating-Protein
recipes %>% 
  ggplot(aes(x=protein, y=rating)) +
  geom_point() +
  ggtitle("Scatterplot of rating against proteins") +
  xlab("Proteins") +
  ylab("Rating") 

# We can observe that the recipes with more than 125 grams of proteins tend to have a higher rating. For instance, few recipes with less than a 3 star rating have more than 125 grams of proteins.


# Scatterplot of Rating-Fat
recipes %>% 
  ggplot(aes(x=fat, y=rating)) +
  geom_point() +
  ggtitle("Scatterplot of rating against fat") +
  xlab("Fat") +
  ylab("Rating") 

# We can observe that the recipes with more than 100 grams of fat tend to have a higher rating. For instance, few recipes with less than a 3 star rating have more than 100 grams of fat.


# Scatterplot of Rating-Sodium
recipes %>% 
  ggplot(aes(x=sodium, y=rating)) +
  geom_point() +
  ggtitle("Scatterplot of rating against sodium") +
  xlab("Sodium") +
  ylab("Rating") 

# We can observe that the recipes with more than 5000 milligrams of sodium tend to have a higher rating. For instance, few recipes with less than a 3 star rating have more than 5000 mg of sodium.

```


#### Correlogram
```{r}
corr_nutritional_values = recipes %>% 
  select(rating, calories, protein, fat, sodium) %>% 
  cor()

corrplot(corr_nutritional_values)

```
The previous scatterplots illuded us that there was somehow a correlation between rating and the nutritional values. This hypothesis has been refuted because the correlation against the rating is almost at zero for all the nutritional values. On the other hand we notice a strong positive correlation between calories and fat as well as between calories and proteins.


#### Grouped Scatter
```{r}

# We decide to plot together the variables which highlight a great level of correlation.

# Grouped scatter of calories and fat
recipes_plot1 <- recipes %>% 
  filter(fat <= 400, calories <= 6000)

recipes_plot1 %>% 
  ggplot(aes(x=calories, y=fat, color=rating)) +
  geom_point() + 
  scale_color_gradientn(colours = rainbow(5))

# Grouped scatter of calories and protein
recipes_plot2 <- recipes %>% 
  filter(protein <= 500, calories <= 6000)

recipes_plot2 %>% 
  ggplot(aes(x=calories, y=protein, color=rating)) +
  geom_point() + 
  scale_color_gradientn(colours = rainbow(5))

# Grouped scatter of protein and fat
recipes_plot3 <- recipes %>% 
  filter(fat <= 400, protein <= 350)

recipes_plot3 %>% 
  ggplot(aes(x=protein, y=fat, color=rating)) +
  geom_point() + 
  scale_color_gradientn(colours = rainbow(5))


# Grouped scatter of protein and sodium
recipes_plot4 <- recipes %>% 
  filter(sodium <= 400, protein <= 350)

recipes_plot4 %>% 
  ggplot(aes(x=protein, y=sodium, color=rating)) +
  geom_point() + 
  scale_color_gradientn(colours = rainbow(5))

```



## Ingredients EDA

### Feature engineering 

```{r Creating ingredients_df}
#Creating a new dataframe with only the ID, title and the ingredients

#through the analysis above, we saw that "drinks" on it's own had only 11 observations, 4 of which also had the value "drink" = 1 --> we decided to merge the two columns to simplify working with a single category called "drink" for all drinks
ingredients_df <- recipes %>% 
  mutate(drink = ifelse(drink == 1 | drinks == 1, 1, 0)) %>% #merging drinks and drink
  select(ID, title, all_of(all_ingredients), rating)

```


#### Adding binary columns

```{r Adding binary columns for each category - Manual Working version}
ingredients_df_bin <- ingredients_df %>%
  mutate(vegetables_bin = as.numeric(if_any(all_of(vegetables_vec), ~.x == 1, na.rm = TRUE)),
         meats_bin = as.numeric(if_any(all_of(all_meats), ~.x == 1, na.rm = TRUE)),
         fish_bin = as.numeric(if_any(all_of(fish_vec), ~.x == 1, na.rm = TRUE)),
         seafood_bin = as.numeric(if_any(all_of(seafood_vec), ~.x == 1, na.rm = TRUE)),
         herbs_bin = as.numeric(if_any(all_of(herbs_vec), ~.x == 1, na.rm = TRUE)),
         nuts_bin = as.numeric(if_any(all_of(nuts_vec), ~.x == 1, na.rm = TRUE)),
         fruits_bin = as.numeric(if_any(all_of(fruits_vec), ~.x == 1, na.rm = TRUE)),
         #drinks_bin = as.numeric(if_any(all_of(drink_no_alcohol_vec), ~.x == 1, na.rm = TRUE)),
         cheese_bin = as.numeric(if_any(all_of(cheeses_vec), ~.x == 1, na.rm = TRUE)),
         dairy_bin = as.numeric(if_any(all_of(dairy_vec), ~.x == 1, na.rm = TRUE)),
         spices_bin = as.numeric(if_any(all_of(spices_vec), ~.x == 1, na.rm = TRUE)),
         cereals_bin = as.numeric(if_any(all_of(cereals_vec), ~.x == 1, na.rm = TRUE)),
         carbs_bin = as.numeric(if_any(all_of(carbs_vec), ~.x == 1, na.rm = TRUE)),
         dessert_bin = as.numeric(if_any(all_of(dessert_vec), ~.x == 1, na.rm = TRUE)),
         egg_bin = (egg)
         ) %>% 
  select(ID, title, contains("bin"), everything())
```

The fact that both select the same number of rows makes having general categories redundant in the dataset. They are not useful to create the binary columns, and they are also not useful to compute the total amount of ingredients in each category per recipe --> let's just not include them in the first place
```{r eval=FALSE}

####testing if I still need to include the general category to create the binary column now that I modified the df to only include recipes with ingredients specified
# 
# #6586
# ingredients_df %>%
#   mutate(vegetables_bin = as.factor(as.numeric(if_any(c(vegetable, all_of(vegetables_vec)), ~.x == 1, na.rm = TRUE)))) %>%
#   filter(vegetables_bin == 1)
# 
# #6586
# ingredients_df %>%
#   mutate(vegetables_bin = as.factor(as.numeric(if_any(all_of(vegetables_vec), ~.x == 1, na.rm = TRUE)))) %>%
#   filter(vegetables_bin == 1)
```


#### Adding total columns

```{r Adding total columns for each category}
ingredients_df_total <- ingredients_df %>%
  mutate(total_ingredients = rowSums(select(., c(all_of(all_ingredients)))),
         total_vegetables = rowSums(select(., c(all_of(vegetables_vec)))),
         total_meat = rowSums(select(., c(all_of(all_meats)))),
         total_fish = rowSums(select(., c(all_of(fish_vec)))),
         total_seafood = rowSums(select(., c(all_of(seafood_vec)))),
         total_herbs = rowSums(select(., c(all_of(herbs_vec)))),
         total_nuts = rowSums(select(., c(all_of(nuts_vec)))),
         total_fruits = rowSums(select(., c(all_of(fruits_vec)))),
         #total_drinks = rowSums(select(., c(all_of(drink_no_alcohol_vec)))),
         total_cheese = rowSums(select(., c(all_of(cheeses_vec)))),
         total_dairy= rowSums(select(., c(all_of(dairy_vec)))),
         total_spices= rowSums(select(., c(all_of(spices_vec)))),
         total_cereals= rowSums(select(., c(all_of(cereals_vec)))),
         total_carbs = rowSums(select(., c(all_of(carbs_vec)))),
         total_dessert = rowSums(select(., c(all_of(dessert_vec))))
         ) %>% 
  select(ID, title, contains("total"), everything())
```

Creating "ingredients_df_full" which contains bin columns, total columns, and original ingredients columns
```{r creating df with bin and total columns}
total_join <- ingredients_df_total %>% 
  select(ID, contains("total"))
  
ingredients_df_full <- ingredients_df_bin %>% 
  left_join(total_join) %>% 
  select(ID, title, rating, contains("bin"), contains("total"), everything())
```

### Analysis part - Within ingredients

#### Frequency of ingredients - binary columns categories

This gives us interesting information about the frequency of each ingredient being present at least once in a recipe. As we can see, there is at least one vegetable in around 6750 recipes out of the 11380 total we have. Inversely, a very low amount of recipes contains at least one type of cereal.
```{r}
#creating a vector with colnames of all the binary columns to be able to select them more easily afterwards
binary_columns <- ingredients_df_bin %>% 
  select(contains("bin")) %>% 
  colnames()

#adding binary columns to ingredients_df
total_categories <- ingredients_df_bin %>% 
  select(ID, all_of(binary_columns)) %>% 
  pivot_longer(-ID, names_to = "category", values_to = "binary_value") %>% 
  group_by(category) %>% 
  summarise(total = sum(binary_value))

#plotting the frequency of binary columns
total_categories %>%
  ggplot(aes(x=reorder(category,total), y=total, fill=total)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(guide = guide_axis(n.dodge=3))+
  scale_fill_viridis() +
  labs(x = "Category", y = "Amount of recipes", title = "Total amount of recipes containing at least one ingredient in defined categories")
```

The below boxplots give us similar information about the amount of recipe which contain at least one ingredient in each category.The only category for which the an ingredient is present at least ince in more than 50% of the recipes is vegetables.
```{r}
ingredients_df_bin %>% 
  select(contains("bin")) %>%
  mutate(across(everything(), as.factor)) %>% 
  plot_bar(order_bar = FALSE)
```


#### Correlation between binary columns, including the rating
We can see that no binary variable correlated to the rating in any way.

We see some somewhat strong negative correlation between vegetables and dessert, and between vegetable and fruits. This makes sense, as these ingredients are rarely found together in recipes. As a side note, we chose to classify tomato as a vegetable and strongly stand by this opinion :)

Concerning positive correlations, we see nuts and desert as highly correlated. This is probably because they go well together in sugary recipes. Additionally, egg and dairy is also correlated at 0.23. This most likely comes from patisserie recipes where eggs and dairy ingredients go hand in hand.
```{r}
ingredients_df_bin %>% 
  select(contains("bin"), rating) %>% 
  plot_correlation()
```
```{r}
#testing if corr matrix changes when we set the binary columns as factor instead of num
ingredients_df_bin %>% 
  select(contains("bin"), rating) %>% 
  mutate(across(-rating, as.factor)) %>% 
  plot_correlation()
#this gives weird results
```
```{r}
#now trying with rating as a factor
ingredients_df_bin %>% 
  select(contains("bin"), rating) %>% 
  mutate(across(rating, as.factor)) %>% 
  plot_correlation()
```


#### Total individual ingredients

```{r}
#Analysis which single ingredient is present in most recipes
df <- ingredients_df %>% 
  select(-title, -rating) %>% 
  pivot_longer(-ID, names_to = "ingredient", values_to = "value")


ing_top10 <- df %>% 
  group_by(ingredient) %>% 
  summarise(total = sum(value)) %>% 
  ungroup() %>% 
  arrange(desc(total)) %>% 
  slice(1:10)

ing_top10 %>% 
  # mutate(ingredient = fct_rev(ingredient)) %>% 
  ggplot(aes(x=reorder(ingredient, total), y=total, fill=ingredient)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis(discrete = TRUE) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  labs(x = "Ingredient", y = "Value", title = "Total amount of recipes containing each ingredient\nTop 10")
```


#### Correlation between number of ingredients and rating

```{r}
ingredients_df_total %>% 
  select(contains("total"), rating) %>% 
  plot_correlation()
```

#### Amount of ingredients per recipe

The number of ingredients per recipe is more or less normally distributed, with a mean around 4.75.
```{r Plotting ingredients totals}
#checking some stuff about the new ingredients table
ingredients_df_total %>% 
  select(ID, title, total_ingredients) %>% 
  ggplot(aes(x=total_ingredients)) + 
  geom_bar() + geom_vline(aes(xintercept=mean(total_ingredients)),color="red", linetype="dashed", size=1)+
   scale_x_continuous(breaks = seq(1, 12, by = 1)) +
  labs(x="Number of ingredients per recipe", y = "Recipe Count", title = "Distrubution of number of ingredients per recipe")
```


```{r test of plot above, eval = FALSE}
#we notice that 117 (no NAs and RAT0, and not duplicated) recipes have 0 ingredients, let's investigate why and how that's possible
ingredients_df_total %>% 
  filter(total_ingredients==0)

#let's pick recipe ID number 1183 which should have poppy and sesame seeds according to the title
recipes %>% 
  filter(ID == 1183) %>% 
  select_if(~ any(. == 1))
#we can see that only 3 variables are equal to 1 here
recipes %>% 
  filter(ID == 365) %>% 
  select_if(~ any(. == 1))

recipes %>% 
  filter(ID == 1089) %>% 
  select_if(~ any(. == 1))

#####
#QUESTION: do we want to keep those recipes?
#####


ingredients_df_total %>% 
  filter(total_ingredients >10)
```
Based on this information, we decide to eliminate those 117 observations which don't contain any ingredients that we have classified in our vectors.
```{r}
ingredients_df_full <- ingredients_df_full %>%
  filter(!total_ingredients == 0)
```


Besides the total amount of ingredients, let's check the amount of ingredients per recipe for the top 3 categories in terms of ingredients frequency (i.e., vegetables, fruits, meats)

```{r Plotting Vegetables total}
ingredients_df_full %>%
  filter(vegetables_bin == 1) %>% 
  select(ID, title, total_vegetables) %>% 
  ggplot(aes(x=total_vegetables)) + 
  scale_x_continuous(breaks = seq(1, 9, by = 1)) +
  geom_bar() + geom_vline(aes(xintercept=mean(total_vegetables)),color="blue", linetype="dashed", size=1)
```


```{r Plotting Fruits total}
ingredients_df_full %>%
  filter(fruits_bin == 1) %>% 
  select(ID, title, total_fruits) %>% 
  ggplot(aes(x=total_fruits)) + 
  scale_x_continuous(breaks = seq(1, 9, by = 1)) +
  geom_bar() + geom_vline(aes(xintercept=mean(total_fruits)),color="blue", linetype="dashed", size=1)
```

```{r Plotting meat totals}
# ingredients_df_full %>% 
#   select(ID, title, total_meat) %>% 
#   ggplot(aes(x=total_meat)) + 
#   geom_bar() + geom_vline(aes(xintercept=mean(total_meat)),color="blue", linetype="dashed", size=1)+
#   labs(x="Number of meats per recipe", y = "Recipe Count", title = "Distrubution of number of meats per recipe")

#let's try to filter by recipes which contain meat to see if my functions work
ingredients_df_full %>% 
  filter(meats_bin == 1) %>% 
  select(ID, title, total_meat) %>% 
  ggplot(aes(x=total_meat)) + 
  geom_bar() + geom_vline(aes(xintercept=mean(total_meat)),color="blue", linetype="dashed", size=1)+
  labs(x="Number of meats per recipe", y = "Recipe Count", title = "Distrubution of number of meats per recipe")

#why do we still have value in 0 meats --> it was because when creating the total meat column in ingredients_df_test I did not include the general meat category
```

### Categories relative to rating

It's hard to interpret the results clearly with so many rating, let's try to reduce it to 2 category --> above 4, and below 4

```{r}
ingredients_df_bin %>% 
  select(contains("bin"), rating) %>%
  mutate(across(everything(), as.factor)) %>% 
  plot_bar(by = "rating", order_bar = FALSE)
```
We not have only 2 categories: recipes with rating above 4 and recipes with ratings below 4. There is no clear relationship in those graphs either, and this confirms the correlation results that we have above.

If we look at vegetables for example, we can see that the proportion of recipes with ratings above 4 is higher for recipes containing no vegetables, when compared to recipes containing at least one vegetable.

```{r}
ingredients_df_bin %>% 
  select(contains("bin"), rating) %>%
  mutate(rating_4 = ifelse(rating>4, 1, 0), across(everything(), as.factor)) %>% 
  select(-rating) %>% 
  plot_bar(by = "rating_4", order_bar = FALSE)
```

## Mixed EDA - ingredients and nutritional value

```{r Left joins to aggregate the data}

recipes_select <- recipes %>% 
  select(ID, title, rating, calories, protein, sodium, fat)

ingredients_select <- ingredients_df_total %>% 
  select(ID, all_of(contains("total")))

recipes_more <- recipes_select %>%
  left_join(ingredients_select, 
           by=c('ID'))

ingredients_bin_select <- ingredients_df_bin %>% 
  select(ID, contains("bin"))


recipes_full <- recipes_more %>%
  left_join(ingredients_bin_select, 
           by=c('ID'))

```

### Correlation between nutritional values and categories
```{r}

recipes_full %>% 
  select(-ID) %>% 
  plot_correlation()


```
For instance we notice a positive correlation between proteins and meats_bin which includes all sorts of meat. Another visible positive correlation is the one between sodium and seafood_bin. We might also want to investigate the relationship between calories and carbs_bin.


### Barplot and boxplot - Meat and Proteins
```{r}

# Barplot
recipes_full %>% 
  ggplot(aes(x = factor(meats_bin), y = protein)) +
  stat_summary(fun.y = mean, geom = "bar") +
  ggtitle("Average amount of proteins per recipe with and without meat") +
  xlab("Presence of Meat or not") +
  ylab("Protein Content in grams")

# Boxplot
recipes_full %>% 
  filter(protein < 450) %>% 
  ggplot(aes(x=factor(meats_bin), y=protein, fill=factor(meats_bin))) +
  geom_boxplot(alpha=0.3) +
  scale_y_continuous(breaks=seq(0,7000,25)) +
  xlab("Presence of Meat or not") +
  ylab("Protein Content in grams") +
  theme(legend.position="none")


# Boxplots per different kinds of meat
recipes_general <- recipes_full %>% 
  select(ID) %>% 
  left_join(recipes, 
           by=c('ID'))

recipes_meat <- recipes_general %>% 
  select(ID, title, rating, calories, protein, fat, sodium, all_of(all_meats))

recipes_meat <- recipes_meat %>% 
  pivot_longer(cols=c("beef", "beef_rib", "beef_shank", "beef_tenderloin", "brisket", "ground_beef", "hamburger", "veal", "bacon", "ham", "pork", "pork_chop", "pork_rib", "pork_tenderloin", "prosciutto", "ground_lamb", "lamb", "lamb_chop", "lamb_shank", "rack_of_lamb", "chicken", "duck", "goose", "poultry", "poultry_sausage", "quail", "turkey", "meatball", "meatloaf", "rabbit", "sausage", "steak", "venison" ),
                    names_to='meats',
                    values_to='yes_or_no') %>% 
  filter(yes_or_no == 1)

recipes_meat %>% 
  filter(protein < 450) %>%
  ggplot(aes(x=meats, y=protein, fill=meats)) +
  geom_boxplot(alpha=0.3) +
  scale_y_continuous(breaks=seq(0,7000,25)) +
  coord_flip() +
  ggtitle("Distribution of proteins per recipe according to different kinds of meat") +
  xlab("Meats") +
  ylab("Protein Content in grams") +
  theme(legend.position="none") 

# Here we want to show which kinds of meat specifically have a high level of proteins

```

### Barplot and boxplot - Seafood and Sodium
```{r}

# Seafood and sodium
recipes_full %>% 
  ggplot(aes(x = factor(seafood_bin), y = sodium)) +
  stat_summary(fun.y = mean, geom = "bar") +
  ggtitle("Average amount of sodium per recipe with and without seafood")  +
  xlab("Presence of Seafood or not") +
  ylab("Sodium Content in milligrams")


# Boxplot
recipes_full %>% 
  filter(sodium < 10000) %>% 
  ggplot(aes(x=factor(seafood_bin), y=sodium, fill=factor(seafood_bin))) +
  geom_boxplot(alpha=0.3) +
  scale_y_continuous(breaks=seq(0,30000,500)) +
  xlab("Presence of Seafood or not") +
  ylab("Sodium Content in milligrams") +
  theme(legend.position="none")



# Boxplots per different kinds of seafood

recipes_seafood <- recipes_general %>% 
  select(ID, title, rating, calories, protein, fat, sodium, all_of(seafood_vec))

recipes_seafood <- recipes_seafood %>% 
  pivot_longer(cols=c("clam", "crab", "lobster", "mussel", "octopus", "oyster", "scallop", "shellfish", "shrimp", "squid" ),
                    names_to='seafoods',
                    values_to='yes_or_no') %>% 
  filter(yes_or_no == 1)

recipes_seafood %>% 
  filter(sodium < 10000) %>%
  ggplot(aes(x=seafoods, y=sodium, fill=seafoods)) +
  geom_boxplot(alpha=0.3) +
  scale_y_continuous(breaks=seq(0,30000,500)) +
  coord_flip() +
  ggtitle("Distribution of sodium per recipe according to different kinds of seafood") +
  xlab("Seafood") +
  ylab("Sodium Content in milligrams") +
  theme(legend.position="none") 

# Here we want to show which kinds of seafood specifically have a high level of sodium

```

### Barplot and boxplot - Carbs and Calories
```{r}

# Carbs and calories
recipes_full %>% 
  ggplot(aes(x = factor(carbs_bin), y = calories)) +
  stat_summary(fun.y = mean, geom = "bar") +
  ggtitle("Average amount of calories per recipe with and without carbohydrates") +
  xlab("Presence of carbohydrates or not") +
  ylab("Calories content")

# Afterwards we would also want to show which kinds of carbs specifically have a high number of calories


# Boxplot
recipes_full %>% 
  ggplot(aes(x=factor(carbs_bin), y=calories, fill=factor(carbs_bin))) +
  geom_boxplot(alpha=0.3) +
  scale_y_continuous(breaks=seq(0,7000,500)) +
  xlab("Presence of Carbs or not") +
  ylab("Calories Content") +
  theme(legend.position="none")


# Boxplots per different kinds of carbs
recipes_carbs <- recipes_general %>% 
  select(ID, title, rating, calories, protein, fat, sodium, all_of(carbs_vec))

recipes_carbs <- recipes_carbs %>% 
  pivot_longer(cols=c("brown_rice", "chickpea", "cornmeal", "couscous", "hominy_cornmeal_masa", "orzo", "pasta", "potato", "rice", "semolina", "sweet_potato_yam", "wild_rice"),
                    names_to='carbs',
                    values_to='yes_or_no') %>% 
  filter(yes_or_no == 1)

recipes_carbs %>% 
  filter(sodium < 10000) %>%
  ggplot(aes(x=carbs, y=calories, fill=carbs)) +
  geom_boxplot(alpha=0.3) +
  scale_y_continuous(breaks=seq(0,7000,500)) +
  coord_flip() +
  ggtitle("Distribution of calories per recipe according to different kinds of food high in carbohydrates") +
  xlab("Carbs") +
  ylab("Calories Content") +
  theme(legend.position="none") 

# Here we want to show which kinds of food high in carbs specifically have a high level of calories
```
## Seasons and Recipe Type EDA

```{r}
#TO ADD WHEN COMPLETED
```
