---
title: "Jordan Test"
output: html_document
date: "2023-04-20"
---

```{r echo=FALSE, message=FALSE}
source(here::here("scripts/setup.R"))
```

# EDA

## Checking for "general" ingredient columns
```{r Meat check}
#checking the meaning of those "general" categories and how they are used
meat_check <- recipes %>% 
  select(meat, all_of(meats_all))

#there is a total of 5232 recipes which contain some type of meat
any_meat <- meat_check %>% 
  filter_if(is.numeric, any_vars(. == 1))

#out of those, 285 recipes where the "meat" category is used
meat_1 <- meat_check%>% 
  filter(meat == 1) %>% 
  select(-meat)


#this allows us to see that in 19 recipes, there is no detailed meat that is specified, the only info we have is that it contains meat
recipes %>%
  select(meat, all_of(meats_all)) %>%
  filter(meat == 1) %>% 
  mutate(sum_meats = rowSums(.)) %>% 
  filter(sum_meats == 1)
#same code to do what I did above
meat_1 %>%
  filter_if(is.numeric, all_vars(. == 0))

#So by not taking into account "meat" we would only lose 19 observations, which aren't specific anyway
```

```{r Veggie Check}
veggie_check <- recipes %>% 
  select(vegetable, all_of(vegetables_vec))

#11222obs with some vegetables_vec in them
veggie_check %>% 
  filter_if(is.numeric, any_vars(. == 1))

#there are 2087 obs with vegetable = 1
legume_1 <- veggie_check %>% 
  filter(vegetable == 1) %>% 
  select(-vegetable)

legume_1 %>%
  filter_if(is.numeric, all_vars(. == 0))
# 290 observations for which vegetable = 1 and not a single vegetable is specified
```

```{r Fish check}
# variable_check <- function(to_check, vec){
#   #10962 obs with some vegetables_vec in them
#   check <- recipes %>% 
#     select(to_check, all_of(vec)) %>% 
#     filter_if(is.numeric, any_vars(. == 1))
#   
#   #there are 69 obs with legume = 1
#   is_1 <- check %>% 
#     filter(to_check == 1) %>% 
#     select(-to_check)
#   
#   # only 4 observations for which legume = 1 and not a single vegetable is specified
#   is_1 %>%
#     filter_if(is.numeric, all_vars(. == 0))
# }

#variable_check("fish", stuff_in_the_water)

#2437 obs with some fish in them
check <- recipes %>% 
  select(fish, seafood, all_of(all_fish_seafood)) %>% 
  filter_if(is.numeric, any_vars(. == 1))

#there are 936 obs with fish = 1
is_1 <- check %>% 
  filter(fish == 1 | seafood == 1) %>% 
  select(-fish, -seafood)

# 455 observations for which fish or seafood = 1 and no specific ingredient is there
is_1 %>%
  filter_if(is.numeric, all_vars(. == 0))
```

```{r Herb check}
#5009 obs with some herb in them
check <- recipes %>% 
  select(herb, all_of(herbs_vec)) %>% 
  filter_if(is.numeric, any_vars(. == 1))

#there are 1681 obs with herb = 1
is_1 <- check %>% 
  filter(herb == 1) %>% 
  select(-herb)

# 1107 observations for which herb = 1 and no specific ingredient is there
is_1 %>%
  filter_if(is.numeric, all_vars(. == 0))
```

```{r Nut check}
#2299 obs with some nut in them
check <- recipes %>% 
  select(nut, all_of(nuts_vec)) %>% 
  filter_if(is.numeric, any_vars(. == 1))

#there are 905 obs with nut = 1
is_1 <- check %>% 
  filter(nut == 1) %>% 
  select(-nut)

# 181 observations for which nut = 1 and no specific ingredient is there
is_1 %>%
  filter_if(is.numeric, all_vars(. == 0))
```

```{r Fruit Check}
#7296 obs with some fruit in them
check <- recipes %>% 
  select(fruit, all_of(fruits_vec)) %>% 
  filter_if(is.numeric, any_vars(. == 1))

#there are 1958 obs with fruit = 1
is_1 <- check %>% 
  filter(fruit == 1) %>% 
  select(-fruit)

# 97 observations for which fruit = 1 and no specific ingredient is there
is_1 %>%
  filter_if(is.numeric, all_vars(. == 0))
```

```{r Cheese check}
#3051 obs with some cheese in them
check <- recipes %>%
  select(cheese, all_of(cheeses_vec)) %>% 
  filter_if(is.numeric, any_vars(. == 1))

#there are 1351 obs with cheese = 1
is_1 <- check %>% 
  filter(cheese == 1) %>% 
  select(-cheese)

# 685 observations for which fruit = 1 and no specific ingredient is there
is_1 %>%
  filter_if(is.numeric, all_vars(. == 0))
```

```{r Dairy check}
#4050 obs with some dairy in them
check <- recipes %>%
  select(dairy, all_of(dairy_vec)) %>% 
  filter_if(is.numeric, any_vars(. == 1))

#there are 1496 obs with dairy = 1
is_1 <- check %>% 
  filter(dairy == 1) %>% 
  select(-dairy)

# 997 observations for which dairy = 1 and no specific ingredient is there
is_1 %>%
  filter_if(is.numeric, all_vars(. == 0))
```

```{r Spice check}
#3530 obs with some spice in them
check <- recipes %>%
  select(spice, all_of(spices_vec)) %>% 
  filter_if(is.numeric, any_vars(. == 1))

#there are 680 obs with spice = 1
is_1 <- check %>%
  filter(spice == 1) %>% 
  select(-spice)

# 455 observations for which spice = 1 and no specific ingredient is there
is_1 %>%
  filter_if(is.numeric, all_vars(. == 0))
```

```{r Drink and Drinks check}
#first checking if we can merge drink and drinks --> 1094 entries with either of those as 1
recipes %>% filter(drink == 1 | drinks == 1)

recipes %>% filter(drink == 1) #1087
recipes %>% filter(drinks == 1) #11

recipes %>% filter(drink == 1 & drinks == 1) #4

#from this we can see that it would make sense to merge the two columns --> done in chunk below


#based on the fact that drinks are both alcoholic and alcohol free, it doesn't really make sense to compare it the the vectors like for the other variables --> we'll just keep "drink"
```

# Feature engineering 

```{r Creating ingredients_df}
#Creating a new dataframe with only the ID, title and the ingredients

#through the analysis above, we saw that "drinks" on it's own had only 11 observations, 4 of which also had the value "drink" = 1 --> we decided to merge the two columns to simplify working with a single category called "drink" for all drinks
ingredients_df <- recipes %>% 
  mutate(drink = ifelse(drink == 1 | drinks == 1, 1, 0)) %>% #merging drinks and drink
  select(ID, title, all_of(general_categories), all_of(all_ingredients))

```

## Adding binary columns 

```{r Testing stuff, eval=FALSE}
#it would be very useful to create a column in recipe in which we have the categories binary to be able to filter more easily

# categories <- recipes %>%
#   select(ID, dairy, all_of(dairy_vec)) %>% 
#   filter_if(is.numeric, any_vars(. == 1)) %>% 
#   mutate(dairy_vec = 1)

test <- ingredients_df %>% 
  mutate(veggies_bin = as.numeric(if_any(all_of(vegetables_vec), ~.x == 1, na.rm = TRUE))) %>% 
  select(all_of(vegetables_vec), veggies_bin)

### testing with paste
name <- (paste("dairy_vec", "bin", sep = "_"))

test <- ingredients_df %>% 
  mutate(name := as.numeric(if_any(c(dairy, all_of(dairy_vec)), ~.x == 1, na.rm = TRUE))) %>% 
  select(all_of(dairy_vec), dairy_vec_bin)

vec <- c("dairy_vec", "vegetables_vec")

for(ingredient in vec){
  new_test <- ingredients_df %>% 
  mutate(dairy_vec_bin = as.numeric(if_any(c(dairy, all_of(dairy_vec)), ~.x == 1, na.rm = TRUE))) %>% 
  select(all_of(dairy_vec), dairy_vec_bin)
}
print(paste("dairy_vec", "bin", sep = "_"))
```

```{r Adding binary columns for each category - Manual Working version}
#doing it manually for, could implement a for loop but not sure I'd gain much time

ingredients_df <- ingredients_df %>%
  #first for categories with a general variable
  mutate(vegetables_bin = as.numeric(if_any(c(vegetable, all_of(vegetables_vec)), ~.x == 1, na.rm = TRUE)),
         meats_bin = as.numeric(if_any(c(meat, all_of(all_meats)), ~.x == 1, na.rm = TRUE)),
         fish_bin = as.numeric(if_any(c(fish, all_of(fish_vec)), ~.x == 1, na.rm = TRUE)),
         seafood_bin = as.numeric(if_any(c(seafood, all_of(seafood_vec)), ~.x == 1, na.rm = TRUE)),
         herbs_bin = as.numeric(if_any(c(herb, all_of(herbs_vec)), ~.x == 1, na.rm = TRUE)),
         nuts_bin = as.numeric(if_any(c(nut, all_of(nuts_vec)), ~.x == 1, na.rm = TRUE)),
         fruits_bin = as.numeric(if_any(c(fruit, all_of(fruits_vec)), ~.x == 1, na.rm = TRUE)),
         drinks_bin = as.numeric(if_any(c(drink, all_of(drink_no_alcohol_vec)), ~.x == 1, na.rm = TRUE)),
         cheese_bin = as.numeric(if_any(c(cheese, all_of(cheeses_vec)), ~.x == 1, na.rm = TRUE)),
         dairy_bin = as.numeric(if_any(c(dairy, all_of(dairy_vec)), ~.x == 1, na.rm = TRUE)),
         spices_bin = as.numeric(if_any(c(spice, all_of(spices_vec)), ~.x == 1, na.rm = TRUE)),
         #now for the other categories which don't have a general variable
         cereals_bin = as.numeric(if_any(all_of(cereals_vec), ~.x == 1, na.rm = TRUE)),
         cereals_bin = as.numeric(if_any(all_of(carbs_vec), ~.x == 1, na.rm = TRUE)),
         dessert_bin = as.numeric(if_any(all_of(dessert_vec), ~.x == 1, na.rm = TRUE)),
         egg_bin = egg
         )
```

```{r For loop template to do the same as above, eval = FALSE}
ingredients_all <- c("dairy", "vegetables_vec", "fruits", "meat", "grains")

for (i in ingredients_all) {
  col_name <- paste0(i, "_bin")
  ingredients_df <- ingredients_df %>%
    mutate(!!col_name := as.numeric(if_any(c(i, all_of(i)), ~.x == 1, na.rm = TRUE)))
}
```

## Adding total columns

```{r Adding total columns for each category}
ingredients_df_test <- ingredients_df %>% 
  mutate(total_ingredients = rowSums(.[,-(1:2)]),
         total_meat = rowSums(select(., c(meat, all_of(all_meats)))),
         total_vegetables = rowSums(select(., c(vegetable, all_of(vegetables_vec))))
         ) %>% 
  select(ID, title, total_ingredients, total_meat, total_vegetables, everything())
```

## Analysis part

### Frequency of ingredients

```{r}
#creating a vector with colnames of all the binary columns to be able to select them more easily afterwards
binary_columns <- ingredients_df %>% 
  select(contains("bin")) %>% 
  colnames()


total_categories <- ingredients_df %>% 
  select(ID, all_of(binary_columns)) %>% 
  pivot_longer(-ID, names_to = "category", values_to = "binary_value") %>% 
  group_by(category) %>% 
  summarise(total = sum(binary_value))

total_categories %>%
  ggplot(aes(x=reorder(category,total), y=total, fill=total)) +
  geom_bar(stat = "identity") +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  scale_fill_viridis() +
  labs(x = "Category", y = "Amount of recipes", title = "Total amount of recipes containing at least one ingredient in defined categories")
```

```{r}
#Analysis which single ingredient comes is present in most recipes
df <- ingredients_df %>% 
  select(-title, -all_of(binary_columns)) %>% 
  pivot_longer(-ID, names_to = "ingredient", values_to = "value")


ing_top10 <- df %>% 
  group_by(ingredient) %>% 
  summarise(total = sum(value)) %>% 
  ungroup() %>% 
  arrange(desc(total)) %>% 
  slice(1:10)

ing_top10 %>% 
  # mutate(ingredient = fct_rev(ingredient)) %>% 
  ggplot(aes(x=reorder(ingredient, total), y=total, fill=ingredient)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis(discrete = TRUE) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  labs(x = "Ingredient", y = "Value", title = "Total amount of recipes containing each ingredient\nTop 10")
```

Here we can see that we have ingredients which in truth are general categories like vegetable, fruit, herb and diary. We don't want this so we will remove those general categories

```{r}
#same as before but removing the general categories first
ing_top10 <- df %>%
  filter(!ingredient %in% general_categories) %>% 
  group_by(ingredient) %>% 
  summarise(total = sum(value)) %>% 
  ungroup() %>% 
  arrange(desc(total)) %>% 
  slice(1:10)

ing_top10 %>% 
  ggplot(aes(x=reorder(ingredient, total), y=total, fill=ingredient)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis(discrete = TRUE) +
  scale_x_discrete(guide = guide_axis(n.dodge=2))+
  labs(x = "Ingredient", y = "Value", title = "Total amount of recipes containing each ingredient\nTop 10 - without general categories")
```

### Amount of ingredients per recipe

```{r Plotting ingredients totals}
#checking some stuff about the new ingredients table
ingredients_df_test %>% 
  select(ID, title, total_ingredients) %>% 
  ggplot(aes(x=total_ingredients)) + 
  geom_bar() + geom_vline(aes(xintercept=mean(total_ingredients)),color="red", linetype="dashed", size=1)

#we notice that 188 recipes have 0 ingredients, let's investigate why and how that's possible
ingredients_df_test %>% 
  filter(total_ingredients==0)

#let's pick recipe ID number 1183 which should have poppy and sesame seeds according to the title
recipes %>% 
  filter(ID == 1183) %>% 
  select_if(~ any(. == 1))
#we can see that only 3 variables are equal to 1 here
recipes %>% 
  filter(ID == 365) %>% 
  select_if(~ any(. == 1))

recipes %>% 
  filter(ID == 1089) %>% 
  select_if(~ any(. == 1))

#####
#QUESTION: do we want to keep those recipes?
#####
```

```{r Plotting meat totals}
ingredients_df_test %>% 
  select(ID, title, total_meat) %>% 
  ggplot(aes(x=total_meat)) + 
  geom_bar() + geom_vline(aes(xintercept=mean(total_meat)),color="blue", linetype="dashed", size=1)

#let's try to filter by recipes which contain meat to see if my functions work
ingredients_df_test %>% 
  filter(meats_bin == 1) %>% 
  select(ID, title, total_meat) %>% 
  ggplot(aes(x=total_meat)) + 
  geom_bar() + geom_vline(aes(xintercept=mean(total_meat)),color="blue", linetype="dashed", size=1)

#why do we still have value in 0 meats --> it was because when creating the total meat column I did not include the general meat category
```

```{r Plotting Vegetables}
ingredients_df_test %>%
  filter(vegetables_bin == 1) %>% 
  select(ID, title, total_vegetables) %>% 
  ggplot(aes(x=total_vegetables)) + 
  scale_x_continuous(breaks = seq(1, 9, by = 1)) +
  geom_bar() + geom_vline(aes(xintercept=mean(total_vegetables)),color="blue", linetype="dashed", size=1)
```

